---
title: "Figure5_plot"
author: "Huan Liu"
date: "2024-06-12"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  eval = FALSE,
  tidy.opts = list(width.cutoff = 120),
  message = TRUE,
  warning = FALSE,
  cache = TRUE,
  label.duplicate = "append"
)

```

# 1 Load packages and datasets

```{r,load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(plot3D)
library(rgl)

library(ggsci)

library(tibble)

library(ggnewscale)

source("./PseudotimeAnalysis.r")

set.seed(12)
theme_set(theme_bw())

load(file = "./Figure5_scRNA_part/Figure5_scRNA_part.Rdata")
#load(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq.Rdata")
#load(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_renamed_sce.Rdata")
#seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```





# 2 Figure5E heatmap
## 2.1 Prepare matrix

```{r}
crv <- SlingshotDataSet(sce)
countMatrix <- sce@assays@data@listData[["counts"]]
counts <- as.matrix(countMatrix)
rm(countMatrix)
```
## 2.2 calculate early driver
```{r, filter low coverage data}
# Assuming your data is in tradeSeq_data
tradeSeq_data <- rowData(sce.trade)$tradeSeq

# 1. Drop the 'Sigma' column
tradeSeq_data <- tradeSeq_data %>% dplyr::select(-Sigma)

# 2. Compute the number of non-missing values per row
num_non_missing <- rowSums(!is.na(tradeSeq_data))

# 3. Identify the rows with the fewest non-missing values
min_non_missing <- min(num_non_missing)
rows_with_fewest_non_missing <- which(num_non_missing == min_non_missing)

# Show the rows with the fewest non-missing values
tradeSeq_data[rows_with_fewest_non_missing, ]

```


Again, we performed the patternTest using the sce.trade object with
abnormal rows excluded.

```{r}
earlyDERes <- earlyDETest(sce.trade[-rows_with_fewest_non_missing,], pairwise = TRUE)
```



```{r,check the headers}
earlyDERes[1:3, ]
```

We want to see which genes

```{r, check top genes}
oPat <- order(earlyDERes$waldStat, 
              decreasing = TRUE)
head(rownames(earlyDERes)[oPat])
```

List the top 10 genes:

```{r,list top 10 genes}
rownames(earlyDERes)[oPat][1:10]
top10_earlyDERes <- rownames(earlyDERes)[oPat][1:10]
write.csv(top10_earlyDERes, file = "./Figure5_scRNA_part/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_EarlyDriver_genes_with_different_pattern_among_all_lineages_20220822.csv", row.names = FALSE )

```

We then plot the top10 early driver genes among the four lineages

```{r,top10 early driver genes among the four lineages}
curvesCols <- c("#619CFF", "#00B9E3","#00BA38","#93AA00")


    for (gene in top10_earlyDERes) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_different_early_driver_among_all_lineages", "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
```

Below, we first write a function to get the top10 genes for a given
comparison, then we implemented it for each pairwise comparison Here we
will pick the top 10 early driver genes differentially activated between
two lineages

```{r, function to get top10 genes in 1vs1 comparision}
# Function to get top 10 genes for a given comparison
getTopEDGenes <- function(column_name) {
    o <- order(earlyDERes[[column_name]], decreasing = TRUE)
    return(names(sce.trade)[o[1:10]])
}

# Apply the function for each pairwise comparison
topEDGenes_1vs2 <- getTopEDGenes("waldStat_1vs2")
topEDGenes_1vs3 <- getTopEDGenes("waldStat_1vs3")
topEDGenes_1vs4 <- getTopEDGenes("waldStat_1vs4")
topEDGenes_2vs3 <- getTopEDGenes("waldStat_2vs3")
topEDGenes_2vs4 <- getTopEDGenes("waldStat_2vs4")
topEDGenes_3vs4 <- getTopEDGenes("waldStat_3vs4")

# Build the resulting dataframe
topEDGenes_1v1 <- data.frame(
    Lineage_1_vs_2 = topEDGenes_1vs2,
    Lineage_1_vs_3 = topEDGenes_1vs3,
    Lineage_1_vs_4 = topEDGenes_1vs4,
    Lineage_2_vs_3 = topEDGenes_2vs3,
    Lineage_2_vs_4 = topEDGenes_2vs4,
    Lineage_3_vs_4 = topEDGenes_3vs4
)

```

```{r,check headers}
head(topEDGenes_1v1)
```

```{r,list top10 genes}
write.csv(topEDGenes_1v1, file = "./Figure5_scRNA_part/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_early_driver_genes_between_two_lineages_20220822.csv", row.names = FALSE )
```

I also want to get the unique genes in each comparison

```{r,unique top genes}
unique_genes <- lapply(1:ncol(topEDGenes_1v1), function(col_idx) {
  current_genes <- topEDGenes_1v1[[col_idx]]
  other_genes <- unlist(topEDGenes_1v1[,-col_idx])
  unique_genes <- current_genes[!(current_genes %in% other_genes)]
  return(unique_genes)
})

names(unique_genes) <- colnames(topEDGenes_1v1)

# Find the maximum number of unique genes for all comparisons
max_length <- max(sapply(unique_genes, length))

# Function to fill a vector with NAs up to max_length
fill_NA <- function(vec) {
  c(vec, rep(NA, max_length - length(vec)))
}

# Apply the fill_NA function to each vector in unique_genes
filled_genes <- lapply(unique_genes, fill_NA)

# Convert the list to a dataframe
unique_genes_df <- as.data.frame(filled_genes)

unique_genes_df
```



Also,I'll loop through the columns of the topEDGenes_1v1 dataframe and
adjust the curvesCols according to the comparison. For this, we'll
create a helper function to determine the appropriate color codes for
each pairwise comparison: We will save the outputs in png formats.

```{r, DE early driver unique genes}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./Figure5_scRNA_part/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}
# Loop through the columns of unique_genes_df and plot each gene
for (comparison in colnames(unique_genes_df)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in unique_genes_df[[comparison]]) {
        
        # Omit NA genes
        if (is.na(gene)) next
        
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_unique_early_driver_gene_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}


```

```{r,DE early driver genes}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./Figure5_scRNA_part/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

# Loop through the columns of topEDGenes_1v1 and plot each gene
for (comparison in colnames(topEDGenes_1v1)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in topEDGenes_1v1[[comparison]]) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_early_driver_gene_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}

```
## 2.3 Bifurcation heatmap for early driver genes among lineage 1 and 3.

However, we want to know if there are any trends cluster along the
bifurcation.Using hybrid heatmap function, we are trying to see if
slingshot+tradeseq could replace BEAM in monocle2. For instance, we want
to check the trends of the top 200 genes in 100 pseudotime intervals.
Immediately after the last part, the oEarly indicate the early driver
genes differentially activated in lineage 1 or lineage 3. Here we want
to compare Lineage 1 vs Lineage 3. And we will repeat some of the
command lines in the last part.

```{r, get oEarly}
oEarly <- order(earlyDERes$waldStat_1vs3, decreasing = TRUE) # get the most differentially activated genes between lineage 1 and lineage 3
head(rownames(earlyDERes)[oEarly])
```

```{r,list top 100 early drivers}
rownames(earlyDERes)[oEarly][1:100] # list the top 100 early driver genes differentially activated between lineage 1 and lineage 3
```

```{r,smooth the model}
yhatSmooth_false <- predictSmooth(sce.trade, gene = rownames(earlyDERes)[oEarly][1:200], #top 200 genes
                                  nPoints = 100, #idealy, I would use 30-100, larger number will make the scale effect not that significant
                                  tidy = FALSE)

```

Since we know that the matrix counts were arranged along the pseudotime,
and we generate the oEarly by the Waldtest between lineage 1 and 3, we
subset and re-arrange the count matrix to generate heatmap using the
hybrid line below:

```{r,fig.height=30, fig.width=20}
heatSmooth <- pheatmap(t(scale(t(yhatSmooth_false[, c(100:1,201:300)]))), # yhatSmooth_false we used nPoints=100, so we choose 1:100 and 201:300 for lineage 1 and 3 respectively
                       cluster_cols = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = TRUE,
                       border_color = FALSE)
ggsave("./Figure5_scRNA_part/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_1_vs3_bifurcation_heatmap.pdf", plot=heatSmooth, width = 10, height = 25)
```


```{r,fig.height=30, fig.width=20}
library(ggsci)

# Assuming yhatSmooth_false is your data matrix
data_matrix <- t(scale(t(yhatSmooth_false[, c(100:1,201:300)])))

# Perform clustering (if not already done)
# Here, I'm assuming hierarchical clustering with cutree for 6 clusters
hc <- hclust(dist(data_matrix))
clusters <- cutree(hc, k = 6)

# Create color annotation
cluster_colors <- setNames(pal_npg("nrc")(6), unique(clusters))

row_annotation <- data.frame(Cluster = factor(clusters, labels = names(cluster_colors)))
row_annotation_colors <- list(Cluster = cluster_colors)

# Generate pheatmap with row annotation
heatSmooth <- pheatmap(data_matrix, 
                       cluster_cols = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = FALSE,
                       border_color = FALSE,
                       annotation_row = row_annotation,
                       annotation_colors = row_annotation_colors)

# Save the plot
ggsave("./Figure5_scRNA_part/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_1_vs3_bifurcation_heatmap_p_test.pdf", plot=heatSmooth, width = 10, height = 25)

```
Export the cut tree results (k=6)
```{r, export cuttree}
row_cluster <- cutree(heatSmooth$tree_row, k = 6)
## order the matrix
newOrder <- yhatSmooth_false[, c(100:1,201:300)][heatSmooth$tree_row$order,]
newOrder <- as.data.frame(newOrder)
newOrder[,ncol(newOrder)+1] = row_cluster[match(rownames(newOrder),names(row_cluster))]
colnames(newOrder)[ncol(newOrder)] = "Cluster"

## export the results
write.table(newOrder, "./Figure5_scRNA_part/JE_combined_epi_SCT_renamed_slingshot_tradeseq_1vs3_DEarly_bifurcation.csv", quote=F, row.names =T, col.names =T, sep ='\t')
```


```{r,check dim of newOrder}
dim(newOrder)
```
## 2.4 Smooth plot for all the genes in each cluster
```{r}
library(tibble)

newOrder <- newOrder %>% rownames_to_column(var ="Gene")
# Filtering and Selecting the Data

filtered_data <- newOrder %>% 
  filter(Cluster == 1) %>% 
  select(Gene,starts_with("lineage1"), Cluster)

# Reshape the data for plotting
data_long <- filtered_data %>% 
  select(-Cluster) %>% 
  gather(key = "Time_Point", value = "Expression_Level", -Gene)

# View the reshaped data
head(data_long)
```
```{r}
# Remove the "lineage1_" prefix and convert to numeric
data_long$Time_Point <- as.numeric(str_replace(data_long$Time_Point, "lineage1_", ""))

# Calculate the average expression level at each time point
average_expression <- data_long %>% 
  group_by(Time_Point) %>% 
  summarise(Average_Expression = mean(Expression_Level))

# Create the plot
ggplot(average_expression, aes(x = Time_Point, y = Average_Expression)) +
  #geom_line() +
  geom_smooth(se = TRUE, method = "loess") +
  theme_minimal() +
  labs(title = "Average Expression Level of Genes in Cluster 1 Over Time (Lineage 1)", 
       x = "Time Point", 
       y = "Average Expression Level")

```

```{r}
# Convert rownames (gene names) to a column
#newOrder <- newOrder %>% rownames_to_column(var = "Gene")

# Create a color palette
cluster_colors <- pal_npg("nrc")(6)
# Initialize variables for y-axis limits
y_limits <- list()
# Define ylim for each cluster across two lineages
y_limits[[1]] <- c(0,56)
y_limits[[2]] <- c(0,10)
y_limits[[3]] <- c(0,10)
y_limits[[4]] <- c(0,45)
y_limits[[5]] <- c(0,40)
y_limits[[6]] <- c(0,55)

# Loop for each cluster and lineage
for(cluster in 1:6) {
  for(lineage in c("lineage1", "lineage3")) {
    # Filtering for the current cluster and selecting the current lineage columns
    filtered_data <- newOrder %>% 
      filter(Cluster == cluster) %>% 
      select(Gene, starts_with(lineage))

    # Reshape the data for plotting
    data_long <- filtered_data %>% 
      gather(key = "Time_Point", value = "Expression_Level", -Gene)

    # Remove the lineage prefix and convert to numeric
    prefix <- paste0(lineage, "_")
    data_long$Time_Point <- as.numeric(str_replace(data_long$Time_Point, prefix, ""))

    # Sort Time_Point in the required order
    if (lineage == "Lineage1") {
      data_long <- data_long %>% arrange(desc(Time_Point))
    } else {
      data_long <- data_long %>% arrange(Time_Point)
    }

    # Calculate the average expression level at each time point
    average_expression <- data_long %>% 
      group_by(Time_Point) %>% 
      summarise(Average_Expression = mean(Expression_Level))

    # Get the y-axis limits for this cluster
    ylims <- y_limits[[cluster]]

    # Create the plot with cluster-specific y-axis limits
    plot <- ggplot(average_expression, aes(x = Time_Point, y = Average_Expression)) +
      geom_smooth(se = TRUE, method = "loess", color = cluster_colors[cluster], fill = cluster_colors[cluster], alpha = 0.4) +
      theme_minimal() +
      labs(title = paste("Average Expression Level of Genes in Cluster", cluster, "Over Time (", lineage, ")", sep = " "), 
           x = "Time Point", 
           y = "Average Expression Level") +
      ylim(ylims[1], ylims[2])  # Apply cluster-specific y-axis limits

    # Print the plot
    print(plot)
  }
}
```

Merge the two lineages into one

```{r,merge teh plots of two lineages into one}
# Create a color palette
cluster_colors <- pal_npg("nrc")(6)

# Define ylim for each cluster across two lineages
y_limits <- list(c(0, 56), c(0, 10), c(0, 10), c(0, 45), c(0, 40), c(0, 55))


# Loop for each cluster
for(cluster in 1:6) {
    # Initialize an empty dataframe to store combined data
    combined_data <- data.frame()

    # Loop through each lineage and combine data
    for(lineage in c("lineage1", "lineage3")) {
        # Filtering and reshaping the data
        filtered_data <- newOrder %>% 
          filter(Cluster == cluster) %>% 
          select(Gene, starts_with(lineage)) %>% 
          gather(key = "Time_Point", value = "Expression_Level", -Gene)

        # Remove the lineage prefix and convert to numeric
        filtered_data$Time_Point <- as.numeric(str_replace(filtered_data$Time_Point, paste0(lineage, "_"), ""))

        # Sort Time_Point in the required order
        if (lineage == "lineage1") {
            filtered_data <- filtered_data %>% arrange(desc(Time_Point))
        } else {
            filtered_data <- filtered_data %>% arrange(Time_Point)
        }

        # Add a column for lineage
        filtered_data$Lineage <- lineage

        # Combine with previous data
        combined_data <- rbind(combined_data, filtered_data)
    }

    # Calculate the average expression level at each time point for each lineage
    average_expression <- combined_data %>% 
      group_by(Time_Point, Lineage) %>% 
      summarise(Average_Expression = mean(Expression_Level))

    # Get the y-axis limits for this cluster
    ylims <- y_limits[[cluster]]

    # Create the plot with cluster-specific y-axis limits
    plot <- ggplot(average_expression, aes(x = Time_Point, y = Average_Expression, group = Lineage)) +
      geom_smooth(aes(linetype = Lineage), se = TRUE, method = "loess", size = 1.2, color = cluster_colors[cluster], fill = alpha(cluster_colors[cluster], 0.4)) +
      scale_linetype_manual(values = c("lineage1" = "solid", "lineage3" = "dashed")) +
      theme_minimal() +
      labs(title = paste("Combined Average Expression Level of Genes in Cluster", cluster, "Over Time"), 
           x = "Time Point", 
           y = "Average Expression Level") +
      ylim(ylims[1], ylims[2]) +
      theme_bw() +  # Use theme_bw
      theme(
      panel.grid.major = element_blank(),  # Turn off major grid lines
      panel.grid.minor = element_blank(),  # Turn off minor grid lines
      legend.position = "right"
  )

    # Print the plot
    print(plot)
    # Save the plot
    file_name <- paste0("./Figure5_scRNA_part/JE_epi_harmony_imputation_slingshot_ED_bifurcation_lineage_1_vs_3_cluster", cluster, ".pdf")
    ggsave(file_name, plot = plot, device = "pdf", width = 11, height = 8.5)

}
```
#3 Figure 4F GO enrichment for different cluster

```{r,load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(plot3D)
library(rgl)
library(clusterProfiler)
library(org.Mm.eg.db)

source("./PseudotimeAnalysis.r")

set.seed(12)
theme_set(theme_bw())

load(file = "./Figure5_scRNA_part/Figure5_scRNA_part.Rdata")
module_gene <- read.csv("./Figure5_scRNA_part/JE_combined_epi_SCT_renamed_slingshot_tradeseq_1vs3_DEarly_bifurcation.csv",sep="\t")
```


```{r, extract genes in each cluster}
###extract gene names in each cluster

module_gene$gene <- rownames(module_gene)

module_gene_list <- module_gene[,c(201,202)]
```




```{r,GO enrichment}
library(clusterProfiler)
library(org.Mm.eg.db)
library(dplyr)

Module_GO <- data.frame()

for (i in unique(module_gene_list$Cluster)) {
  
  data <- filter(module_gene_list, module_gene_list$Cluster == i)
  df <- bitr(data$gene, 
             fromType = "SYMBOL",
             toType = c("ENTREZID"), 
             OrgDb = "org.Mm.eg.db")  # Convert SYMBOL to ENTREZID
  
  go <- enrichGO(gene = unique(df$ENTREZID),
                 OrgDb = org.Mm.eg.db,
                 keyType = 'ENTREZID',
                 ont = "BP",
                 pAdjustMethod = "BH",
                 pvalueCutoff = 0.05,
                 qvalueCutoff = 0.05,
                 readable = TRUE)
  go_res <- go@result
  
  if (dim(go_res)[1] != 0) {
    go_res$cluster <- i
    Module_GO <- rbind(Module_GO, go_res)
  }
}

# Filter significant terms
Module_GO <- Module_GO[which(Module_GO$qvalue <= 0.05),]
#Module_GO <- Module_GO[, c("ID", "Description", "qvalue", "cluster", "geneID")]
write.csv(Module_GO, file = './Figure5_scRNA_part/Module_GO.csv')

```

#4 Smooth plot for genes of interest
##4.1 Trp63 smooth plot
```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Trp63", curvesCols = curvesCols,size=1/10,
              border = FALSE) +  xlim(0,15) +ylim(0,6)+
  scale_color_manual(values = curvesCols)
ggsave("./Figure5_scRNA_part/Plot_Trp63_lineage1vs3.pdf",  width = 15, height = 5)
```
##4.2 Runx1 smooth plot
```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Runx1", curvesCols = curvesCols,size=1/10,
              border = FALSE) +  xlim(0,15) +ylim(0,6)+
  scale_color_manual(values = curvesCols)
ggsave("./Figure5_scRNA_part/Plot_Runx1_lineage1vs3.pdf",  width = 15, height = 5)
```



#5 Figure 4G tri genes umap
Prepare packages and datasets
```{r,load packages}

library(ggplot2)
library(scales)
library(ggnewscale)
seu_list <- SplitObject(seu, split.by = "orig.ident")
seu_D0 <-seu_list[["D0"]]
```

```{r, extract expression matrix}
#Extract expression matrix from（scaled）
exp <- seu_D0@assays[["MAGIC_SCT"]]@data
exp <- as.data.frame(exp)#With row for genes, column for cell
#extract UMAP coordinates
position=seu@reductions$umap@cell.embeddings
position <- as.data.frame(position)
position$cell <- rownames(position)
```

## Trp63
```{r, extract gene expression for genes of interest,fig.width=8,fig.height=7}

gene <- c("Trp63", "Odam", "Lor") # Blue color for Odam, Green color for Lor

data_gene <- position
for (i in 1:length(gene)) {
  gene_exp <- exp[gene[i],]
  gene_exp <- t(gene_exp)
  gene_exp <- as.data.frame(gene_exp)
  gene_exp$cell <- rownames(gene_exp)
  data_gene <- merge(data_gene, gene_exp, by='cell')
}

colnames(data_gene)
rownames(data_gene) <- data_gene$cell
data_gene <- data_gene[, -1]


threshold <- 0.5


data_Trp63 <- data_gene[, c(1:3)]
data_Trp63 <- data_Trp63[which(data_Trp63$Trp63 > threshold),]

data_Odam <- data_gene[, c(1, 2, 4)]
data_Odam <- data_Odam[which(data_Odam$Odam > threshold),]

data_Lor <- data_gene[, c(1, 2, 5)]
data_Lor <- data_Lor[which(data_Lor$Lor > threshold),]

# 设置绘图参数
point_size <- 1.0
point_alpha <- 0.6

ggplot() +
  new_scale("color") +
  geom_point(data = data_Trp63, aes(x = UMAP_1, y = UMAP_2, colour = Trp63), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "red") +
  new_scale("color") +
  geom_point(data = data_Odam, aes(x = UMAP_1, y = UMAP_2, colour = Odam), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#619CFF") +
  new_scale("color") +
  geom_point(data = data_Lor, aes(x = UMAP_1, y = UMAP_2, colour = Lor), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#00BA38") +
  geom_point(data = data_gene, aes(x = UMAP_1, y = UMAP_2), size = 0.5, shape = 16, color = 'lightgrey', alpha = 0.1) +
    theme_minimal() +
  theme(
    panel.grid = element_blank(),  # remove grid
    panel.border = element_rect(color = "black", fill = NA, size = 1)  # add black border
  )
ggsave("./Figure5_scRNA_part/JE_K5_epi_Trp63_umap.pdf")
```

## Klf4
```{r, extract gene expression for genes of interest,fig.width=8,fig.height=7}

gene <- c("Klf4", "Odam", "Lor") # Blue color for Odam, Green color for Lor

data_gene <- position
for (i in 1:length(gene)) {
  gene_exp <- exp[gene[i],]
  gene_exp <- t(gene_exp)
  gene_exp <- as.data.frame(gene_exp)
  gene_exp$cell <- rownames(gene_exp)
  data_gene <- merge(data_gene, gene_exp, by='cell')
}

colnames(data_gene)
rownames(data_gene) <- data_gene$cell
data_gene <- data_gene[, -1]


threshold <- 0.5


data_Trp63 <- data_gene[, c(1:3)]
data_Trp63 <- data_Trp63[which(data_Trp63$Klf4 > threshold),]

data_Odam <- data_gene[, c(1, 2, 4)]
data_Odam <- data_Odam[which(data_Odam$Odam > threshold),]

data_Lor <- data_gene[, c(1, 2, 5)]
data_Lor <- data_Lor[which(data_Lor$Lor > threshold),]

# 设置绘图参数
point_size <- 1.0
point_alpha <- 0.6

ggplot() +
  new_scale("color") +
  geom_point(data = data_Trp63, aes(x = UMAP_1, y = UMAP_2, colour = Klf4), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "red") +
  new_scale("color") +
  geom_point(data = data_Odam, aes(x = UMAP_1, y = UMAP_2, colour = Odam), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#619CFF") +
  new_scale("color") +
  geom_point(data = data_Lor, aes(x = UMAP_1, y = UMAP_2, colour = Lor), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#00BA38") +
  geom_point(data = data_gene, aes(x = UMAP_1, y = UMAP_2), size = 0.5, shape = 16, color = 'lightgrey', alpha = 0.1) +
    theme_minimal() +
  theme(
    panel.grid = element_blank(),  
    panel.border = element_rect(color = "black", fill = NA, size = 1)  
  )
ggsave("./Figure5_scRNA_part/JE_K5_epi_Klf4_umap.pdf")
```

## Itga3
```{r, extract gene expression for genes of interest,fig.width=8,fig.height=7}

gene <- c("Itga3", "Odam", "Lor") # Blue color for Odam, Green color for Lor

data_gene <- position
for (i in 1:length(gene)) {
  gene_exp <- exp[gene[i],]
  gene_exp <- t(gene_exp)
  gene_exp <- as.data.frame(gene_exp)
  gene_exp$cell <- rownames(gene_exp)
  data_gene <- merge(data_gene, gene_exp, by='cell')
}

colnames(data_gene)
rownames(data_gene) <- data_gene$cell
data_gene <- data_gene[, -1]


threshold <- 0.5


data_Trp63 <- data_gene[, c(1:3)]
data_Trp63 <- data_Trp63[which(data_Trp63$Itga3 > 0),]

data_Odam <- data_gene[, c(1, 2, 4)]
data_Odam <- data_Odam[which(data_Odam$Odam > threshold),]

data_Lor <- data_gene[, c(1, 2, 5)]
data_Lor <- data_Lor[which(data_Lor$Lor > threshold),]


point_size <- 1.0
point_alpha <- 0.6

ggplot() +
  new_scale("color") +
  geom_point(data = data_Trp63, aes(x = UMAP_1, y = UMAP_2, colour = Itga3), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "red") +
  new_scale("color") +
  geom_point(data = data_Odam, aes(x = UMAP_1, y = UMAP_2, colour = Odam), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#619CFF") +
  new_scale("color") +
  geom_point(data = data_Lor, aes(x = UMAP_1, y = UMAP_2, colour = Lor), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#00BA38") +
  geom_point(data = data_gene, aes(x = UMAP_1, y = UMAP_2), size = 0.5, shape = 16, color = 'lightgrey', alpha = 0.1) +
    theme_minimal() +
  theme(
    panel.grid = element_blank(),  
    panel.border = element_rect(color = "black", fill = NA, size = 1)  
  )
ggsave("./Figure5_scRNA_part/JE_K5_epi_Itga3_umap.pdf")
```

## Itgb1
```{r, extract gene expression for genes of interest,fig.width=8,fig.height=7}

gene <- c("Itgb1", "Odam", "Lor") # Blue color for Odam, Green color for Lor

data_gene <- position
for (i in 1:length(gene)) {
  gene_exp <- exp[gene[i],]
  gene_exp <- t(gene_exp)
  gene_exp <- as.data.frame(gene_exp)
  gene_exp$cell <- rownames(gene_exp)
  data_gene <- merge(data_gene, gene_exp, by='cell')
}

colnames(data_gene)
rownames(data_gene) <- data_gene$cell
data_gene <- data_gene[, -1]


threshold <- 0.5


data_Trp63 <- data_gene[, c(1:3)]
data_Trp63 <- data_Trp63[which(data_Trp63$Itgb1 > 0.2),]

data_Odam <- data_gene[, c(1, 2, 4)]
data_Odam <- data_Odam[which(data_Odam$Odam > threshold),]

data_Lor <- data_gene[, c(1, 2, 5)]
data_Lor <- data_Lor[which(data_Lor$Lor > threshold),]


point_size <- 1.0
point_alpha <- 0.6

ggplot() +
  new_scale("color") +
  geom_point(data = data_Trp63, aes(x = UMAP_1, y = UMAP_2, colour = Itgb1), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "red") +
  new_scale("color") +
  geom_point(data = data_Odam, aes(x = UMAP_1, y = UMAP_2, colour = Odam), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#619CFF") +
  new_scale("color") +
  geom_point(data = data_Lor, aes(x = UMAP_1, y = UMAP_2, colour = Lor), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#00BA38") +
  geom_point(data = data_gene, aes(x = UMAP_1, y = UMAP_2), size = 0.5, shape = 16, color = 'lightgrey', alpha = 0.1) +
    theme_minimal() +
  theme(
    panel.grid = element_blank(), 
    panel.border = element_rect(color = "black", fill = NA, size = 1) 
  )
ggsave("./Figure5_scRNA_part/JE_K5_epi_Itgb1_umap.pdf")
```

## Runx1
```{r, extract gene expression for genes of interest,fig.width=8,fig.height=7}

gene <- c("Runx1", "Odam", "Lor") # Blue color for Odam, Green color for Lor

data_gene <- position
for (i in 1:length(gene)) {
  gene_exp <- exp[gene[i],]
  gene_exp <- t(gene_exp)
  gene_exp <- as.data.frame(gene_exp)
  gene_exp$cell <- rownames(gene_exp)
  data_gene <- merge(data_gene, gene_exp, by='cell')
}

colnames(data_gene)
rownames(data_gene) <- data_gene$cell
data_gene <- data_gene[, -1]


threshold <- 0.5

data_Trp63 <- data_gene[, c(1:3)]
data_Trp63 <- data_Trp63[which(data_Trp63$Runx1 > 0.2),]

data_Odam <- data_gene[, c(1, 2, 4)]
data_Odam <- data_Odam[which(data_Odam$Odam > threshold),]

data_Lor <- data_gene[, c(1, 2, 5)]
data_Lor <- data_Lor[which(data_Lor$Lor > threshold),]

# 设置绘图参数
point_size <- 1.0
point_alpha <- 0.6

ggplot() +
  new_scale("color") +
  geom_point(data = data_Trp63, aes(x = UMAP_1, y = UMAP_2, colour = Runx1), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "red") +
  new_scale("color") +
  geom_point(data = data_Odam, aes(x = UMAP_1, y = UMAP_2, colour = Odam), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#619CFF") +
  new_scale("color") +
  geom_point(data = data_Lor, aes(x = UMAP_1, y = UMAP_2, colour = Lor), shape = 16, size = point_size, alpha = point_alpha) +
  scale_colour_gradient(low = "lightgrey", high = "#00BA38") +
  geom_point(data = data_gene, aes(x = UMAP_1, y = UMAP_2), size = 0.5, shape = 16, color = 'lightgrey', alpha = 0.1) +
    theme_minimal() +
  theme(
    panel.grid = element_blank(),  
    panel.border = element_rect(color = "black", fill = NA, size = 1) 
  )
ggsave("./Figure5_scRNA_part/JE_K5_epi_Runx1_umap.pdf")
```



#X save Rdata



```{r}

save(sce_dm, #sce in diffusion map
     sce, #JE_combined_epi slingshot object
     counts, # count matrix derived from sce
     seu, # Seuratobject for D0-D5
     seu_D0, # Seuratobject for D0
     sce.trade, #sce fitGAM
     earlyDERes, # lineage specific driver genes between lineages
     startRes,#progenitor marker genes within all lineages
     startRes_all, #progenitor marker genes within each lineages
     customRes, # lineage specific genes within a specific pseudotime point of a specific lineage
     endRes, # differentiated cell type markers between lineages
     patternRes, ##  genes with different expression patterns
     module_gene, ## different cluster of genes in bifurcation heatmap
     Module_GO, ## GO annotation for different cluster of genes in bifurcation heatmap
     file = "./Figure5_scRNA_part/Figure5_scRNA_part.Rdata")
```

```{r}
gc()
rm(list=ls())
```





