---
title: "Gingival Palatal Epithelium ATAC"
author: "Huan"
date: "2023-05-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE,
  cache = TRUE
)
```



# 1 Environment setup
```{r}
gc()
library("DiffBind")
library("ggplot2")
library("pheatmap")
library("dplyr")
set.seed(12)
theme_set(theme_bw())
```


# 2 Load data and sample list
```{r,load data and sample list}
atac <- read.csv("Gingiva_palate_ATAC_samplelist.csv")
names(atac)
atac <- dba(sampleSheet="Gingiva_palate_ATAC_samplelist.csv")
atac
plot(atac)
```

# 3 Count reads
```{r}
atac_all <- dba.count(atac, summit=250)
atac_all #examine the new DBA object

```

Here we want to check the reads that overlap a consensus peak
```{r}

info <- dba.show(atac_all)
libsizes <- cbind(LibReads=info$Reads, FRiP=info$FRiP, PeakReads=round(info$Reads * info$FRiP))
rownames(libsizes) <- info$ID
libsizes
```


Raw plot for centered counts
```{r}
plot(atac_all)
```

# 4 Normalization
```{r}
## here we consider the default normalization for our example, obtained using the dba.normalize function:
atac_norm <- dba.normalize(atac_all) #By default, the data are normalized based on sequencing depth.
## Here we check the normalizing factors:
norm <- dba.normalize(atac_norm, bRetrieve=TRUE)
norm
```

```{r}
## The default library-size based methods results in all the library sizes being normalized to be the same (the mean library size):
normlibs <- cbind(FullLibSize=norm$lib.sizes, NormFacs=norm$norm.factors, NormLibSize=round(norm$lib.sizes/norm$norm.factors))
rownames(normlibs) <- info$ID
normlibs
```

Raw plot for centered and normalized counts
```{r}
plot(atac_norm)
```


# 5 Differential analysis
## 5.1 Setup contrast 
```{r}
atac_con <- dba.contrast (atac_norm,  categories=DBA_TISSUE, minMembers = 2)
atac_con
dba.show(atac_con, bContrasts=TRUE)
```

## 5.2 Differential analysis
```{r}
atac_diff <- dba.analyze(atac_con) 

dba.show(atac_diff, bContrasts=TRUE)
atac_diff
```

Heatmap for the correlation
```{r}
plot(atac_diff, contrast=1)
```

## 5.3 Plot of differential bindings
```{r, fig.width=10, fig.height=5}
dba.plotPCA(atac_diff, label=DBA_TISSUE)
```

```{r, fig.width=10, fig.height=8}
for (i in 1:length(atac_diff[["contrasts"]]) ){
  dba.plotMA(atac_diff, contrast=i, bNormalized=FALSE)
}

```

```{r,fig.height=6, fig.width=4}
for (j in 1:length(atac_diff[["contrasts"]]) ){
  dba.plotHeatmap(atac_diff,contrast = j, correlations = FALSE)
}


```

## 5.3 Save DBA object for further use
```{r}
saveRDS(atac_diff, file = "Gingiva_Palate_epi_NFR_Deseq2_240505.rds")
```


#6 Export Diff Peak
```{r}
diffPeaks <- dba.report(atac_diff,th = 1) %>%
  data.frame() %>%
  mutate(type = case_when(Fold >= 1 & FDR < 0.05 ~ "sigUp",
                          Fold <= -1 & FDR < 0.05 ~ "sigDown",
                          .default = "nonSig"))
```


check 
```{r}
table(diffPeaks$type)
```

Export all diff
```{r}
write.csv(diffPeaks, file = "./Deseq2/Gingiva_Palate_epi_NFR_Deseq2_diffAll_240505.csv",row.names = F)
```

# 7 Annotate diff peaks
## 7.1 Load packages

```{r}
library(ChIPseeker)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
```

## 7.2 Annotation for Palate enriched NFR (sigUp)
```{r}
sigdiff <- diffPeaks %>% dplyr::filter(type == "sigUp") %>%
  GRanges()

peakAnno <- annotatePeak(sigdiff, tssRegion = c(-3000, 3000),
                         TxDb = TxDb.Mmusculus.UCSC.mm10.knownGene)

peakAnno_df <- data.frame(peakAnno)

# id transformation
ids <- mapIds(org.Mm.eg.db, keys = peakAnno_df$geneId, keytype = "ENTREZID", column = "SYMBOL")
peakAnno_df$GeneName <- ids

# output
write.csv(peakAnno_df, 'Deseq2/diffbind_palate_sigPeak_anno.csv', row.names = FALSE)

# plot
plotAnnoBar(peakAnno)
ggsave(filename = "./Deseq2/diffbind_palate_sigPeak_anno.pdf")
```
## 7.2 Annotation for Junctional enriched NFR (sigDown)
```{r}
sigdiff <- diffPeaks %>% dplyr::filter(type == "sigDown") %>%
  GRanges()

peakAnno <- annotatePeak(sigdiff, tssRegion = c(-3000, 3000),
                         TxDb = TxDb.Mmusculus.UCSC.mm10.knownGene)

peakAnno_df <- data.frame(peakAnno)

# id transformation
ids <- mapIds(org.Mm.eg.db, keys = peakAnno_df$geneId, keytype = "ENTREZID", column = "SYMBOL")
peakAnno_df$GeneName <- ids

# output
write.csv(peakAnno_df, 'Deseq2/diffbind_Junctional_sigPeak_anno.csv', row.names = FALSE)

# plot
plotAnnoBar(peakAnno)
ggsave(filename = "./Deseq2/diffbind_Junctional_sigPeak_anno.pdf")
```
# 8Export diffPeak
```{r}
diffPeaks <- read.csv("./Deseq2/Gingiva_Palate_epi_NFR_Deseq2_diffAll_240505.csv")

types <- c("sigUp","sigDown","nonSig")

# x = 2
lapply(seq_along(types),function(x){
  tmp <- diffPeaks %>% dplyr::filter(type == types[x]) %>%
    dplyr::select(seqnames,start,end)

  write.table(tmp,file = paste0("./Deseq2/",types[x],".bed"),
              quote = F,col.names = F,row.names = F,sep = "\t")
})
```

