---
title: "Mouse JE repair scRNA harmony epi imputation pseudotime with slingshot with diffusion map approach"
author: "Huan"
date: "1/4/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  eval = FALSE,
  tidy.opts = list(width.cutoff = 120),
  message = TRUE,
  warning = FALSE,
  cache = TRUE,
  label.duplicate = "append"
)

```

This note was used for pseudotime imputation using
`JE_combined_epi_harmony_MAGIC_renamed.rds`. We will also work with the
two main paths `JE_combined_epi_MAGIC_GE7_20221128.rds` and
`JE_combined_epi_MAGIC_GE4_20221128.rds` We employed the following
approaches: slingshot Root cells were determined using scVelo generated
from
`Working with scVelo for integrated junctional epithelium (D0, D3 and D5) scRNA-seq (dynamic).ipynb`

Here we will show how to transfer the embedding information generated
from Seurat to monocle analysis

# 1 Slingshot for scRNA data

## 1.0 Load data and packages

```{r,load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(rgl)
set.seed(12)
theme_set(theme_bw())


seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
Idents(seu) <- c("celltype")
DefaultAssay(seu) <- c("SCT")
```


```{r}
seu
```
```{r}
DimPlot(seu, label = T, pt.size = 1)+NoLegend()
```



refer to
`https://bustools.github.io/BUS_notebooks_R/slingshot.html#introduction`
Unfortunately, slingshot does not natively support ggplot2. So this is a
function that assigns colors to each cell in base R graphics.

```{r}
#' Assign a color to each cell based on some value
#' 
#' @param cell_vars Vector indicating the value of a variable associated with cells.
#' @param pal_fun Palette function that returns a vector of hex colors, whose
#' argument is the length of such a vector.
#' @param ... Extra arguments for pal_fun.
#' @return A vector of hex colors with one entry for each cell.
cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}
```

```{r}
cell_colors <- cell_pal(seu$celltype, brewer_pal("qual", "Set2"))
```

```{r}
cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )
```

## 1.1 Generate slingshot object using seurat

```{r}
seu.sc <- as.SingleCellExperiment(seu)
table(seu.sc$ident)

```

```{r}
save(seu.sc, #intitial sce object after conversion
     cell_colors, # default color
     cb_palette, ##  color for different assigned celltype
     file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_renamed_sce.Rdata")
```


Resume
```{r}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(plot3D)
library(rgl)
set.seed(12)
theme_set(theme_bw())

load(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_renamed_sce.Rdata")
seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```

```{r}
source("./PseudotimeAnalysis.r")
```

From our previous notebook using scVelo, we identified GE1 as the start, with GE3, GE4 and GE7 as end points.

```{r}
sce  <- slingshot(seu.sc, clusterLabels = seu.sc$ident, reducedDim = 'UMAP',
                  start.clus = c("GE1"), 
                  end.clus = c("GE3", "GE4", "GE7"),
                  approx_points = 150)               
```
You may use our previous color apporoach using `cb_palette`.

```{r}
SlingshotDataSet(sce)
plot(reducedDims(sce)$UMAP,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.2, label = T)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

## 1.2 Probing with Diffusion map

```{r}
library(destiny)

curve(find_dm_k(n), 0, 5624, xname = 'n')
```
```{r}
set.seed(123)
find_sigmas(sce@int_colData@listData[["reducedDims"]]@listData[["HARMONY"]])
```
```{r}

run_diffMap <- function(data=data, condition=condition,sigma="local", k = 20)
  {
  destinyObj <- as.ExpressionSet(as.data.frame(t(data))) 
  destinyObj$condition <- factor(condition) 
  dm <- DiffusionMap(destinyObj, sigma, k) 
  return(dm)

}

plot_eigenVal <- function(dm=dm){ 
  linepad <- .5 
  plot( 
    eigenvalues(dm), 
    ylim = 0:1,
    pch = 20,
    xlab ='Diffusion component (DC)', 
    ylab ='Eigenvalue'
  )
}

plot_dm_3D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, size=0.5, legend = NULL){
  cond <- factor(condition) 
  col <- factor(condition) 
  levels(col) <- colours 
  col <- as.vector(col) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame(
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]] 
    ) 
  colnames(data) <- DCs
  plot3d( 
    data, 
    bg=col, 
    col=col, 
    size=size, 
    box = FALSE 
    )
  if (is.null(legend)==FALSE){ 
    legend3d("topright", legend = unique(condition), pch = 21, pt.bg = unique(col), cex=1.5, inset=c(0.02), bty = "n") 
     }
}

plot_dm_3D_in_2D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, outline.color = "grey20", size=1, pch = 21, theta = 40, phi = 40, bty ="b"){
  cond <- factor(condition) 
  cols <- factor(condition) 
  levels(cols) <- colours 
  cols <- as.vector(cols) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame( 
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]]
    )
  colnames(data) <- DCs
  if(pch == 21){
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              bg = cols,
              col = outline.color, 
              pch=pch,
              cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta =theta, phi = phi, bty = bty)
    } else {
      scatter3D(x=as.matrix(data[,DCs[1]]),
                y=as.matrix(data[,DCs[2]]), 
                z=as.matrix(data[,DCs[3]]),
                colvar =as.numeric(factor(condition)), col = colours,
                pch=pch, cex = size,
                xlab = DCs[1],
                ylab = DCs[2],
                zlab = DCs[3],
                theta =theta,
                phi = phi,
                bty = bty)
    }
}



get_pseudotime <- function(pseudotime, wthres=wthres){
    pseudoT <- list()
    for(lineage in 1:length(pseudotime@metadata[["curves"]]))local({
        curve <- pseudotime@metadata[["curves"]][[lineage]]
        lambda <- curve$lambda
        weight <- curve$w
        ps <- curve$lambda
        ps[weight < wthres] <- NA
        ps <- rankKeepNA(ps)
        pseudoT[[lineage]] <<- ps
    })
    df <- t(do.call("rbind",pseudoT))
    colnames(df) <- names(pseudotime@metadata[["curves"]])
    return(df)
}

```


```{r}
sce_dm <- run_diffMap(t(seu@reductions$harmony@cell.embeddings), colData(sce)$ident, sigma = 29.8, k = 1000) 
plot_eigenVal(dm = sce_dm)
```
```{r, fig.height=20, fig.width=20}
my_color<-cb_palette
names(my_color) <- levels(seu)
#plot.DiffusionMap(sce_dm,dims = 1:2,col_by = phenoData(dataset(sce_dm)))
```
```{r}
sce_lineage <- slingshot(sce_dm@eigenvectors[,c(1:4)], clusterLabels =factor(colData(sce)$ident), 
                         start.clus = 'GE1',
                         end.clus = c("GE3", "GE4", "GE7"),
                         allow.breaks=FALSE, maxit = 1000, shrink.method ="density", thresh = 0.001, extend = "n")
```

```{r}
SlingshotDataSet(sce_lineage)
```


```{r, fig.height=20, fig.width=20}
splom(~sce_dm@eigenvectors[, 1:5], groups =sce$ident, col = my_color, main ="sce_lineage",key = list(space="right", 
                                                                                                      points = list(pch = 19, col =my_color), 
                                                                                                      text = list(c(levels(sce$ident)))))

ggsave()
```




```{r}
plot_dm_3D(dm = sce_dm, dc = c(1,2,3), condition = seu@active.ident, colour = my_color, size = 3) 
rglwidget()
rgl.postscript('./OutputFigure/3dplot.pdf', fmt = 'pdf')
```


Get pseudotime with slingshot
```{r}
#Get pseudotime with slingshot
sce_pseudotime <- get_pseudotime(sce_lineage, wthres = 0.90)
rownames(sce_pseudotime) <- colnames(x = sce)

lineage1 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 1]), 1])
lineage2 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 2]), 2])
lineage3 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 3]), 3])
lineage4 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 4]), 4])
lineage5 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 5]), 5])

```

```{r}
length(lineage1)
```
```{r}
curves <- slingCurves(sce_lineage)

```


## 1.3 Visualization 


```{r, fig.height=10, fig.width=10}
plot_dm_3D_in_2D(dm = sce_dm, dc = c(1, 2, 3), condition = seu@active.ident, 
                 colour = my_color, size = 1, theta = 90, phi = 150, pch = 21, bty = "g")
```





Highlight each lineage
```{r}
highlight_cells_per_lineage_dmplot3D_2 <- function(object, dm=dm, dc=c(1:3), highlight.cells, condition=condition, colours=colours, lineage.curve=NULL, size=1, pch = 21, outline.color = "#E5E4E2", bg.color = "#E5E4E2", title="Diffusion map",phi = 40, theta = 40, add.curve=NULL, bty ="b"){
    
    # Adjusting the transparency for bg.color and outline.color
    bg.color <- grDevices::adjustcolor(bg.color, alpha.f = 0.7)
    outline.color <- grDevices::adjustcolor(outline.color, alpha.f = 0.2)
    
    cell.use.s <- highlight.cells
    
    cols <- factor(condition)
    levels(cols) <- colours
    cols <- as.vector(cols)
    DCs <- paste("DC",dc, sep="")
    
    data <- data.frame(
    dm@eigenvectors[,DCs[1]],
    dm@eigenvectors[,DCs[2]],
    dm@eigenvectors[,DCs[3]],
    Cell=row.names(object@reductions$pca@cell.embeddings),
    Ident=condition,
    Color=cols
    )
    colnames(data) <- c(DCs,"Cell","ident","Color")
    
    data2 = subset(data, data$Cell %in% cell.use.s)
    
    scatter3D(x=as.matrix(data[,DCs[1]]), 
    y=as.matrix(data[,DCs[2]]), 
    z=as.matrix(data[,DCs[3]]), 
    bg = bg.color, 
    col = outline.color, 
    pch=pch, 
    cex = size, 
    xlab = DCs[1], 
    ylab = DCs[2], 
    zlab = DCs[3], 
    theta =theta, 
    phi = phi, main =title, bty = bty)
    
    scatter3D(x=as.matrix(data2[,DCs[1]]), y=as.matrix(data2[,DCs[2]]), z=as.matrix(data2[,DCs[3]]), bg = as.matrix(data2[,"Color"]), 
    col = outline.color, pch=pch, cex = size, xlab = DCs[1], ylab = DCs[2], zlab = DCs[3], theta =theta, phi = phi, add=TRUE, bty = bty)
    
    if(!is.null(add.curve)){
        c <- add.curve
        scatter3D(x = c$s[c$ord,DCs[1]], y =c$s[c$ord,DCs[2]], z = c$s[c$ord,DCs[3]], col = "black", type = "l", ticktype = "detailed", lwd = 4, add = TRUE, xlab = DCs[1], ylab = DCs[2], zlab = DCs[3], theta =theta, phi = phi, bty = bty)
    }
    
}

```



```{r}
glist <- list()
celllist <- list(names(lineage1), names(lineage2), names(lineage3), names(lineage4), names(lineage5))
for (i in 1:5){
	# Start the png device
  filename <- paste0("./OutputFigure/JE_1st_impute_slingshot_phi40_Branch_", i, ".png")
  png(filename, width=1800, height=1800)
  cell.use.s <- celllist[[i]]
  title <- paste("Branch", i, sep="")
  g <- highlight_cells_per_lineage_dmplot3D_2(object = seu, dm = sce_dm, dc = c(1, 2, 3), 
                                            condition = seu@active.ident, colours = my_color,
                                            highlight.cells = cell.use.s, size=0.5,theta = 90,  phi = 10, bty = "n", title = title)
  g <- g + scatter3D(x = curves$Lineage1$s[curves$Lineage1$ord, 1], 
          y = curves$Lineage1$s[curves$Lineage1$ord, 2], 
          z = curves$Lineage1$s[curves$Lineage1$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage2$s[curves$Lineage2$ord, 1], 
          y = curves$Lineage2$s[curves$Lineage2$ord, 2], 
          z = curves$Lineage2$s[curves$Lineage2$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage3$s[curves$Lineage3$ord, 1], 
          y = curves$Lineage3$s[curves$Lineage3$ord, 2], 
          z = curves$Lineage3$s[curves$Lineage3$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage4$s[curves$Lineage4$ord, 1], 
          y = curves$Lineage4$s[curves$Lineage4$ord, 2], 
          z = curves$Lineage4$s[curves$Lineage4$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage5$s[curves$Lineage5$ord, 1], 
          y = curves$Lineage5$s[curves$Lineage5$ord, 2], 
          z = curves$Lineage5$s[curves$Lineage5$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)

  glist[[i]] <- g
    # Close the graphics device, saving the plot
  dev.off()
}
```

```{r}

highlight_cells_per_lineage_UMAPplot <- function(object, highlight.cells, pt.size = 2, pt.shape = 21, outline.color = "grey30", bg.color = "#d3d3d3", hi.color, outline.size = 0.2,title="Umap"){
    
    cell.use.s <- highlight.cells
    
    data.plot=data.frame(object@reductions$umap@cell.embeddings,row.names(object@reductions$pca@cell.embeddings),object@active.ident)
    colnames(data.plot) <- c("UMAP_1","UMAP_2","Cell","ident")
    
    g=ggplot(data.plot,aes(x=UMAP_1,y=UMAP_2))
    g=g+geom_point(size=pt.size, shape=pt.shape, color=outline.color, fill=bg.color, stroke = outline.size)
    g=g+geom_point(data=subset(data.plot, data.plot$Cell %in% cell.use.s), aes(fill =ident), color=outline.color, size=pt.size, shape=pt.shape, stroke = outline.size)
    g=g+scale_fill_manual(values = hi.color,name="")+ggtitle(title)
    g=g+theme_grey()+
    theme(
    axis.text=element_text(size=16),
    axis.title=element_text(size=16),
    legend.text = element_text(size =16),
    legend.title = element_text(size =16 ,face="bold"),
    plot.title = element_text(size=18, face="bold", hjust = 0.5),
    aspect.ratio=1
    )
    return(g)
}

```



```{r}
glist <- list()
celllist <- list(names(lineage1), names(lineage2),names(lineage3), names(lineage4),names(lineage5))
for (i in 1:5){
  cell.use.s <- celllist[[i]]
  title <- paste("Branch", i, sep="")
  g <- highlight_cells_per_lineage_UMAPplot(seu, highlight.cells = cell.use.s, hi.color = my_color, title = title)
  glist[[i]] <- g
}
```
```{r, fig.width=20, fig.height=20}
plot_grid(glist[[1]], glist[[2]], glist[[3]], glist[[4]], glist[[5]], ncol = 3)
ggsave(filename = "./OutputFigure/pJE_1st_impute_slingshot_lineage_umap.png", width = 30, height = 30)
```







## 1.4 Initial plot for all Lineages in UMAP

By default, you can use cell_colors for celltype color options

```{r}
plot(reducedDims(sce)$UMAP,  col = cell_colors, pch=16, asp = 1, cex=0.5)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

```{r}
plot(reducedDims(sce)$PCA,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.5)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

You may also use our previous color apporoach using `cb_palette`.

```{r}
plot(reducedDims(sce)$UMAP,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.2)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

```{r}
plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2)                         
lines(SlingshotDataSet(sce), lwd=2,  col='black')
```

You may set alpha (from scales package). With bigger `cex` and smaller
`alpha`.

```{r}
plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette[sce$celltype],0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)                         
lines(SlingshotDataSet(sce), lwd=2,  col='black')
```

Obviously, we will use UMAP for the downstream plot Below, we will
present how to "hide" lineage you don't want to show.

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = i)
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}



```

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = i)
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black')
}



```

When running `SlingshotDataSet(sce)` we will see the composite of
different lineages: lineages: 4 Lineage1: GE1 GE8 GE5 GE9 GE7\
Lineage2: GE1 GE8 GE2 GE6\
Lineage3: GE1 GE8 GE4\
Lineage4: GE1 GE8 GE3

## 1.5 Pseudotime plot for Lienage 1

For instance, if you are especially interested in GE7 related, ie,
Lineage 1, you may run the following commands for plot

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage1"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage1")
lines(SlingshotDataSet(sce), linInd = 1)



```

```{r}

pt <- slingPseudotime(sce)

nms <- colnames(pt)

pal <- viridis(100, end = 0.95)
par(mfrow = c(1, 2))

colors <- pal[cut(pt[,"Lineage1"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage1", ylim=c(-10,10))
lines(SlingshotDataSet(sce), linInd = 1)



plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2,ylim=c(-10,10))  
lines(SlingshotDataSet(sce), linInd = 1, type="l")
```

From the above plot, we can clearly see the path for Lineage 1: GE1 GE8
GE5 GE9 GE7\
If you don't want to see the color for other unwanted cluster, you may
try this "dumb" approach, by coloring other cluster as gray (#808080)

```{r}
cb_palette_Lineage1 <- c("#F8766D", #GE1
                "#C0C0C0", #GE2
                "#C0C0C0", #GE3
                "#C0C0C0", #GE4
                "#00C19F", #GE5
                "#C0C0C0", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3"  #GE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage1[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 1, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage1[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 1)
```

## 1.6 Pseudotime plot for other lineages

For instance, if you are especially interested in GE3 related, ie,
Lineage 4, you may run the following commands for plot

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage4"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage4")
lines(SlingshotDataSet(sce), 
      linInd = 4 #change to 4 if you are working with Lineage4
      )




```

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage4"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage4")
lines(SlingshotDataSet(sce), linInd = 4)


plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2)  
lines(SlingshotDataSet(sce), linInd = 4, type="l")
```

From the above plot, we can clearly see the path for Lineage4: GE1 GE8
GE3\
If you don't want to see the color for other unwanted cluster, you may
try this "dumb" approach, by coloring other cluster as gray (#808080)

```{r}

cb_palette_Lineage4 <- c("#F8766D", #GE1
                "#C0C0C0", #GE2
                "#93AA00", #GE3
                "#C0C0C0", #GE4
                "#C0C0C0", #GE5
                "#C0C0C0", #GE6
                "#C0C0C0", #GE7
                "#DB72FB", #GE8
                "#C0C0C0" #GE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage4[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 4, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage4[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 4)
```

Lineage2: GE1 GE8 GE2 GE6

```{r}

cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )

cb_palette_Lineage2 <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#C0C0C0", #GE3
                "#C0C0C0", #GE4
                "#C0C0C0", #GE5
                "#00B9E3", #GE6
                "#C0C0C0", #GE7
                "#DB72FB", #GE8
                "#C0C0C0" #GE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage2[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 2, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage2[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 2)
```
From the above plot Lineage 2 is not a real one, in such resolution

Lineage3: GE1 GE8 GE4

```{r}

cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )

cb_palette_Lineage3 <- c("#F8766D", #JE1
                "#C0C0C0", #JE2
                "#C0C0C0", #JE3
                "#00BA38", #JE4
                "#C0C0C0", #JE5
                "#C0C0C0", #JE6
                "#C0C0C0", #JE7
                "#DB72FB", #JE8
                "#C0C0C0" #JE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage3[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 3, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage3[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 3)
```

## 1.7 Save Outputs

```{r,save output}
saveRDS(sce, file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot.rds")

```
