---
title: "Figure_2_plots"
author: "Huan Liu"
date: "2024-07-03"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  eval = FALSE,
  tidy.opts = list(width.cutoff = 120),
  message = TRUE,
  warning = FALSE,
  cache = TRUE,
  label.duplicate = "append"
)

```

# 1 Load packages and datasets

```{r,load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(plot3D)
library(rgl)
library(RColorBrewer)

source("./diffprop_functions.R")

set.seed(12)
theme_set(theme_bw())

load(file = "./Figure2_scRNA_part/Figure2_scRNA_part.Rdata")

#sce <-readRDS(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot.rds")
#JE_combined_epi <-readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")

#JE_K5 <- readRDS(file = "./SeuratObject/JE_K5_harmony_celltype_20240708.rds")
#JE_K5_epi <- readRDS(file = "./SeuratObject/JE_K5_epi_MAGIC_harmony_renamed.rds")
#JE_K5_MAGIC <- readRDS(file = "./SeuratObject/JE_K5_harmony_MAGIC_20240710.rds")


```

#2 Figure 2B Diffusion map
From our previous notebook using scVelo, we identified GE1 as the start, with GE3, GE4 and GE7 as end points.
##2.1 Slingshot
```{r, generate slingshot object}
DefaultAssay(JE_combined_epi) <- c("SCT")
JE_combined_epi_sce  <- slingshot(JE_combined_epi, clusterLabels = JE_combined_epi$ident, reducedDim = 'UMAP',
                  start.clus = c("GE1"), 
                  end.clus = c("GE3", "GE4", "GE7"),
                  approx_points = 150)               
```
You may use our previous color apporoach using `cb_palette`.

```{r, check slingshot object}
SlingshotDataSet(JE_combined_epi_sce)
plot(reducedDims(JE_combined_epi_sce)$UMAP,  col =cb_palette[JE_combined_epi_sce$celltype], pch=16, asp = 1, cex=0.2, label = T)                       
lines(SlingshotDataSet(JE_combined_epi_sce), lwd=2, type = 'lineage', col='black')
```

## 2.2 Probing with Diffusion map
In order to probe for the best routes of trajectories, we employed diffusion map.

```{r, set basic parameters for diffusion map}
library(destiny)

curve(find_dm_k(n), 0, 5624, xname = 'n')
my_color<-cb_palette
names(my_color) <- levels(JE_combined_epi)
set.seed(123)
find_sigmas(JE_combined_epi_sce@int_colData@listData[["reducedDims"]]@listData[["HARMONY"]])
```

```{r, modify function for probeing and ploting}

run_diffMap <- function(data=data, condition=condition,sigma="local", k = 20)
  {
  destinyObj <- as.ExpressionSet(as.data.frame(t(data))) 
  destinyObj$condition <- factor(condition) 
  dm <- DiffusionMap(destinyObj, sigma, k) 
  return(dm)

}

plot_eigenVal <- function(dm=dm){ 
  linepad <- .5 
  plot( 
    eigenvalues(dm), 
    ylim = 0:1,
    pch = 20,
    xlab ='Diffusion component (DC)', 
    ylab ='Eigenvalue'
  )
}

plot_dm_3D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, size=0.5, legend = NULL){
  cond <- factor(condition) 
  col <- factor(condition) 
  levels(col) <- colours 
  col <- as.vector(col) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame(
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]] 
    ) 
  colnames(data) <- DCs
  plot3d( 
    data, 
    bg=col, 
    col=col, 
    size=size, 
    box = FALSE 
    )
  if (is.null(legend)==FALSE){ 
    legend3d("topright", legend = unique(condition), pch = 21, pt.bg = unique(col), cex=1.5, inset=c(0.02), bty = "n") 
     }
}

plot_dm_3D_in_2D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, outline.color = "grey20", size=1, pch = 21, theta = 40, phi = 40, bty ="b"){
  cond <- factor(condition) 
  cols <- factor(condition) 
  levels(cols) <- colours 
  cols <- as.vector(cols) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame( 
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]]
    )
  colnames(data) <- DCs
  if(pch == 21){
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              bg = cols,
              col = outline.color, 
              pch=pch,
              cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta =theta, phi = phi, bty = bty)
    } else {
      scatter3D(x=as.matrix(data[,DCs[1]]),
                y=as.matrix(data[,DCs[2]]), 
                z=as.matrix(data[,DCs[3]]),
                colvar =as.numeric(factor(condition)), col = colours,
                pch=pch, cex = size,
                xlab = DCs[1],
                ylab = DCs[2],
                zlab = DCs[3],
                theta =theta,
                phi = phi,
                bty = bty)
    }
}


plot_dm_3D_in_2D_2 <- function(dm, dc=c(1:3), condition, colours, outline.color = "grey20", size=1, pch = 21, theta = 40, phi = 40, bty ="b", alpha=1){
  
  cond <- factor(condition) 
  cols <- factor(condition) 
  levels(cols) <- colours 
  cols <- as.vector(cols) 
  
  # Adjust colors for transparency
  cols <- grDevices::adjustcolor(cols, alpha=alpha)
  outline.color <- grDevices::adjustcolor(outline.color, alpha=alpha)
  colours <- grDevices::adjustcolor(colours, alpha=alpha)
  
  DCs <- paste("DC",dc, sep="")
  data <- data.frame( 
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]]
  )
  colnames(data) <- DCs
  
  if(pch == 21){
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              bg = cols,
              col = outline.color, 
              pch=pch,
              cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta = theta, phi = phi, bty = bty)
  } else {
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              colvar = as.numeric(factor(condition)), col = colours,
              pch=pch, cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta = theta,
              phi = phi,
              bty = bty)
  }
}



get_pJE_combined_epidotime <- function(pJE_combined_epidotime, wthres=wthres){
    pJE_combined_epidoT <- list()
    for(lineage in 1:length(pJE_combined_epidotime@metadata[["curves"]]))local({
        curve <- pJE_combined_epidotime@metadata[["curves"]][[lineage]]
        lambda <- curve$lambda
        weight <- curve$w
        ps <- curve$lambda
        ps[weight < wthres] <- NA
        ps <- rankKeepNA(ps)
        pJE_combined_epidoT[[lineage]] <<- ps
    })
    df <- t(do.call("rbind",pJE_combined_epidoT))
    colnames(df) <- names(pJE_combined_epidotime@metadata[["curves"]])
    return(df)
}

```


```{r, gernate diffusion map}
JE_combined_epi_sce_dm <- run_diffMap(t(JE_combined_epi@reductions$harmony@cell.embeddings), colData(JE_combined_epi_sce)$ident, sigma = 29.8, k = 1000) 
plot_eigenVal(dm = JE_combined_epi_sce_dm)
```

```{r, regernate slingshot using different eigenvectors}
JE_combined_epi_sce_lineage <- slingshot(JE_combined_epi_sce_dm@eigenvectors[,c(1:4)], clusterLabels =factor(colData(JE_combined_epi_sce)$ident), 
                         start.clus = 'GE1',
                         end.clus = c("GE3", "GE4", "GE7"),
                         allow.breaks=FALSE, maxit = 1000, shrink.method ="density", thresh = 0.001, extend = "n")
```

```{r, check slingshot objects generated using diffusion map}
SlingshotDataSet(JE_combined_epi_sce_lineage)
```


```{r, check diffusion map in different dimension, fig.height=20, fig.width=20}
splom(~JE_combined_epi_sce_dm@eigenvectors[, 1:5], groups =JE_combined_epi_sce$ident, col = my_color, main ="JE_combined_epi_sce_lineage",key = list(space="right", 
                                                                                                      points = list(pch = 19, col =my_color), 
                                                                                                      text = list(c(levels(JE_combined_epi_sce$ident)))))

ggsave(filename = "Figure2_scRNA_part/splom_dm_JE_combined_epi_sce_dm.png")
```
Since the end points are GE3, GE4 and GE7, we chose dc1, dc2 and dc3
```{r,plot 3d version of dm}
plot_dm_3D(dm = JE_combined_epi_sce_dm, dc = c(1,2,3), condition = JE_combined_epi@active.ident, colour = my_color, size = 3) 
rglwidget()
rgl.snapshot('./Figure2_scRNA_part/Figure2B_3dplot.png', fmt = "png") # this is not working. Thus, after adjusting the angles, I save the rgl plot into PNG, and modifed in Adobe illustrator.
```


#3 Figure2C Krt14 and Krt5 smooth plot

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Krt14", curvesCols = curvesCols,size=1/10,
              border = FALSE) +  xlim(0,15) +ylim(0,6)+
  scale_color_manual(values = curvesCols)
ggsave("./Figure2_scRNA_part/Plot_Krt14_lineage1vs3.pdf",  width = 15, height = 5)
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Krt5", curvesCols = curvesCols,size=1/10,
              border = FALSE) + xlim(0,15) +ylim(0,6)+
  scale_color_manual(values = curvesCols)
ggsave("./Figure2_scRNA_part/Plot_Krt5_lineage1vs3.pdf",  width = 15, height = 5)
```
#4 Summary of JE_K5

## 4.1 UMAP for celltype
```{r, initial plots}
Idents(JE_K5_MAGIC) <- JE_K5_MAGIC$celltype
#group_by_cluster
plot1 = DimPlot(JE_K5_MAGIC, reduction = "umap", pt.size=0.01, label = T, repel=T)
# group_by_sample
plot2 = DimPlot(JE_K5_MAGIC, reduction = "umap", pt.size=0.01, group.by = "orig.ident")
plot3 = DimPlot(JE_K5_MAGIC, reduction = "umap", pt.size=0.01, label = T, split.by  = "orig.ident")
# combine
plotc <- plot1 + plot2
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_umap.pdf", plot = plotc, width = 20, height = 10)
plot3
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_umap_split.pdf", plot = plot3, width = 24, height = 10)

DimPlot(JE_K5_MAGIC, reduction = "umap", pt.size=0.01, label = T)
DimPlot(JE_K5_MAGIC, reduction = "umap", pt.size=0.01, label = T,group.by = "orig.ident")

```
## 4.2 Dotplot for marker genes in JE_1st
```{r, setup gene list}
#markers refer to https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7359078/ and marker genes in each cluster
genes_to_check <- c("Col1a1", "Postn","Nfia", #Dermal
                    "Epcam","Krt14","Krt24","Krt5","Mki67","Krtdap","Atf3",  #Epithelium
                    "Ccl3","Cxcl2", #Macrophage
                    "Lyz2","Cd74", #Monocyte
                    "Plvap","Egfl7",#Endothelial cell
                    "Icos","Cxcr6", #T cell
                    "Mpz","Plp1", #Glial cell
                    "Rgs5","Myh11", #Pericyte,
                    "Ccl21a","Mmrn1", #Lymphatic endothelial cell
                    "Hbb-bt","Hba-a1", #Erythroid cell
                    "Cd79a","Cd79b" #B cell
                             )
```

```{r,dotplot for marker genes}

# Define the desired order for celltype
desired_order <- c("Fibroblast","Basal","Cycling","Intermidate","Keratinocyte","JE","Macrophage","Monocyte",
                   "Endothelial cell","T Cell", "Glial cell","Pericyte","Lymphatic Vessel","B cell")
desired_order <- rev(desired_order)

# Set the identities in the Seurat object to celltype
JE_K5$celltype <- factor(JE_K5$celltype, levels = desired_order)
Idents(JE_K5) <- JE_K5$celltype

# Generate the dot plot
DotPlot(JE_K5, features = unique(genes_to_check)) +
  theme_bw() +
  theme(panel.grid = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5),
        panel.border = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_blank()) +
  labs(x = NULL, y = NULL) + 
  guides(size = guide_legend(order = 3)) +
  scale_color_gradientn(values = seq(0, 1, 0.2), colours = c('#295B9F', '#FFFFFF', '#D23B3E'))

# Save the dot plot
ggsave("./Figure2_scRNA_part/JE_K5_harmony_dotplot_feature.pdf", width = 8.5, height = 5)
```


## 4.3 Celltype percentage display
```{r}
# Load necessary libraries
library(Seurat)
library(data.table)
library(ggplot2)
library(reshape2)
library(dplyr)

# Extract meta data
md <- JE_K5_MAGIC@meta.data %>% as.data.table
md  # The resulting md object has one "row" per cell

# Count the number of cells per unique combinations of "orig.ident" and "celltype"
cell_counts <- md[, .N, by = c("orig.ident", "celltype")]

# Cast the data from long to wide format
split_count <- dcast(cell_counts, orig.ident ~ celltype, value.var = "N", fill = 0)

# Melt the data back to long format for ggplot2
split_count_melt <- melt(split_count, id.vars = "orig.ident", variable.name = "celltype", value.name = "Number")

# Save the cell count data to a CSV file
write.csv(split_count_melt, file = "./Figure2_scRNA_part/JE_K5_MAGIC_harmony_cell_number_celltype.csv", row.names = FALSE)

# Define the desired order for celltype
desired_order <- c("Fibroblast","Basal","Cycling","Intermidate","Keratinocyte","JE","Macrophage","Monocyte",
                   "Endothelial cell","T Cell", "Glial cell","Pericyte","Lymphatic Vessel","B cell")


# Convert celltype to factor with the desired order
split_count_melt$celltype <- factor(split_count_melt$celltype, levels = desired_order)

# Generate a stacked bar plot with a white theme and black border for unnormalized data
ggplot(split_count_melt, aes(fill = orig.ident, x = celltype, y = Number)) +
  geom_bar(position = "fill", stat = "identity") +
  ggtitle("Percentage of cells in each cluster") +
  xlab("celltype") +
  ylab("Percentage of cells") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5),
        panel.border = element_rect(color = "black", fill = NA, size = 0.5))

ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_harmony_Percentage_cell_in_each_celltype_unnormalized.pdf", height = 3, width = 7.5)

# Normalize cell counts to the total number in each sample
split_count$total <- rowSums(split_count[, -1])

# Calculate the percentage for each cluster
split_count_percent <- split_count %>%
  mutate(across(-c(orig.ident, total), ~ . / total * 100, .names = "{.col}_percent"))

# Reshape data for ggplot2
split_count_percent_melt <- melt(split_count_percent, id.vars = "orig.ident", 
                                 variable.name = "celltype", 
                                 value.name = "Percentage",
                                 measure.vars = grep("_percent$", names(split_count_percent), value = TRUE))

# Remove "_percent" from celltype
split_count_percent_melt$celltype <- gsub("_percent", "", split_count_percent_melt$celltype)

# Filter out rows with NA in celltype
split_count_percent_melt <- split_count_percent_melt[!is.na(split_count_percent_melt$celltype), ]

# Convert celltype to factor with the desired order
split_count_percent_melt$celltype <- factor(split_count_percent_melt$celltype, levels = desired_order)

# Save the percentage data to a CSV file
write.csv(split_count_percent_melt, file = "./Figure2_scRNA_part/JE_K5_MAGIC_harmony_cell_percentage_celltype.csv", row.names = FALSE)

# Generate a stacked bar plot with a white theme and black border for normalized data
ggplot(split_count_percent_melt, aes(fill = orig.ident, x = celltype, y = Percentage)) +
  geom_bar(position = "fill", stat = "identity") +
  ggtitle("Percentage of cells in each celltype") +
  xlab("celltype") +
  ylab("Percentage of cells") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_harmony_Percentage_cell_in_each_celltype_normalized.pdf", height = 3, width = 7.5)

# Generate a stacked bar plot for percentage of cells (alternative layout)
ggplot(split_count_percent_melt, aes(x = celltype, y = Percentage)) +
  geom_col(aes(fill = orig.ident), color = "black") + 
  theme_bw() + 
  theme(panel.grid = element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_harmony_Percentage_cell_in_each_celltype_normalized_stack.pdf", height = 3, width = 7.5)

# Generate a dodged bar plot for percentage of cells
ggplot(split_count_percent_melt, aes(x = celltype, y = Percentage)) +
  geom_col(aes(fill = orig.ident), color = "black", position = "dodge") + 
  theme_bw() + 
  theme(panel.grid = element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_harmony_Percentage_cell_in_each_celltype_normalized_dodge.pdf", height = 3, width = 7.5)

```
## 4.4 Heatmap for celltype percentage in each sample
```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)

# Define the desired order for celltype
desired_order <- c("Fibroblast","Basal","Cycling","Intermidate","Keratinocyte","JE","Macrophage","Monocyte",
                   "Endothelial cell","T Cell", "Glial cell","Pericyte","Lymphatic Vessel","B cell")
                   
desired_order <- rev(desired_order)



# Extract meta data
md <- JE_K5_MAGIC@meta.data %>% as.data.table
md  # The resulting md object has one "row" per cell

# Count the number of cells per unique combinations of "orig.ident" and "celltype"
cell_counts <- md[, .N, by = c("orig.ident", "celltype")]

# Calculate the total number of cells in each sample (orig.ident)
cell_totals <- cell_counts[, .(total = sum(N)), by = orig.ident]

# Merge the total counts with the cell counts
cell_counts <- merge(cell_counts, cell_totals, by = "orig.ident")

# Calculate the percentage of cells in each celltype within each orig.ident
cell_counts[, Percentage := (N / total) * 100]

# Convert celltype to factor with the desired order
cell_counts$celltype <- factor(cell_counts$celltype, levels = desired_order)

# Generate the heatmap plot
ggplot(cell_counts, aes(x = orig.ident, y = celltype, fill = Percentage)) +
  geom_tile(color = NA) + # Remove the color outline in each tile
  geom_text(aes(label = sprintf("%.1f", Percentage)), color = "black", size = 3) +
  scale_fill_gradient2(low = "#0AC94E", mid = "#FFEEA6", high = "#DD4D46", midpoint = 20, limits = c(0, 40)) +
  labs(title = "Percentage of celltype in each sample",
       x = "Sample Day", y = "Cluster Renamed") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.border = element_rect(color = "black", fill = NA, size = 1))

# Save the heatmap plot
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_harmony_heatmap_cell_proportion.pdf", height = 5, width = 7)

```


# 5 Feature plot
## 5.1 Whole
1) GFP lor odam
```{r,GFP lor odam whole,fig.width=15,fig.height=15}
FeaturePlot(JE_K5, 
                  features = c("GFP","Lor","Odam"),
                  cols = c("grey85", brewer.pal(9,"YlOrRd")),
                  ncol=2, 
                  pt.size = 0.1,
                  min.cutoff = 0.1,
                  max.cutoff =5,
            label=T,repel = T)
ggsave("./Figure2_scRNA_part/JE_K5_GFP_Lor_Odam.pdf")
```
```{r,GFP lor odam whole magic,fig.width=15,fig.height=15}
DefaultAssay(JE_K5_MAGIC) <- "MAGIC_SCT"
FeaturePlot(JE_K5_MAGIC, 
                  features = c("GFP","Lor","Odam"),
                  cols = c("grey85", brewer.pal(9,"YlOrRd")),
                  ncol=2, 
                  pt.size = 0.1,
                  min.cutoff = 0.1,
                  max.cutoff =5,
            label=T,repel = T)
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_GFP_Lor_Odam.pdf")
```



##5.2 Seperate

```{r,GFP lor odam split MAGIC,fig.width=25,fig.height=27}
DefaultAssay(JE_K5_MAGIC) <- "MAGIC_SCT"
FeaturePlot(JE_K5_MAGIC, 
                  features = c("GFP","Krt5","Lor","Odam"),
                  cols = c("grey85", brewer.pal(9,"YlOrRd")),
                  ncol=2, split.by  = "orig.ident",
                  pt.size = 0.1,
                  min.cutoff = 1,
                  max.cutoff =5,
            label=T,repel = T)
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_GFP_Krt5_Lor_Odam_split_1.pdf")
```
```{r,GFP lor odam split MAGIC,fig.width=25,fig.height=27}
DefaultAssay(JE_K5_MAGIC) <- "MAGIC_SCT"
FeaturePlot(JE_K5_MAGIC, 
                  features = c("GFP","Krt5","Lor","Odam"),
                  cols = c("grey85", brewer.pal(9,"YlOrRd")),
                  ncol=2, split.by  = "orig.ident",
                  pt.size = 0.1,
                  min.cutoff = 0.1,
                  max.cutoff =5,
            label=T,repel = T)
ggsave("./Figure2_scRNA_part/JE_K5_MAGIC_GFP_Krt5_Lor_Odam_split_0.1.pdf")
```

#6 GFP percentage in epi 
Here we use violin plot to show GFP+ cells distribution in different celltype between D0 and D5
```{r}

JE_K5_epi_D0_D5 <- subset(JE_K5_epi, orig.ident !="JE_K5_D3") 
```



```{r}
Idents(JE_K5_epi_D0_D5) <- "celltype"
custom_col = c("#F8766D", "#00BA38")
VlnPlot(JE_K5_epi_D0_D5, 
        features = c("GFP"),
        pt.size = 0,
        ncol = 1,split.by = "orig.ident",
        cols = custom_col)
ggsave("./Figure2_scRNA_part/JE_K5_epi_D0_D5_GFP_vlnplot.pdf", height = 4, width = 7)
```










#X save Rdata

```{r}
save(JE_combined, #seurat object fro combined with celltype information
     JE_combined_epi, #seurat object for combined_epi with celltype informationsce_dm, #sce in diffusion map
     JE_K5, # JE_K5 merged data
     JE_K5_epi,# JE_K5_epi population imputated with MAGIC 
     JE_K5_MAGIC, #JE_K5 imputation
     sce, #JE_combined_epi slingshot object
     counts, # count matrix derived from sce
     sce.trade, #sce fitGAM

     file = "./Figure2_scRNA_part/Figure2_scRNA_part.Rdata")
```

# Before shut down Rserver
```{r}
gc()
rm(list=ls())
```

