---
title: "Mouse JE repair scRNA harmony epi imputation pseudotime with slingshot with diffusion map approach"
author: "Huan"
date: "1/4/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  eval = FALSE,
  tidy.opts = list(width.cutoff = 120),
  message = TRUE,
  warning = FALSE,
  cache = TRUE,
  label.duplicate = "append"
)

```

This note was used for pseudotime imputation using
`JE_combined_epi_harmony_MAGIC_renamed.rds`. We will also work with the
two main paths `JE_combined_epi_MAGIC_GE7_20221128.rds` and
`JE_combined_epi_MAGIC_GE4_20221128.rds` We employed the following
approaches: slingshot Root cells were determined using scVelo generated
from
`Working with scVelo for integrated junctional epithelium (D0, D3 and D5) scRNA-seq (dynamic).ipynb`

Here we will show how to transfer the embedding information generated
from Seurat to monocle analysis

# 1 Slingshot for scRNA data

## 1.0 Load data and packages

```{r,load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(rgl)
set.seed(12)
theme_set(theme_bw())


seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
Idents(seu) <- c("celltype")
DefaultAssay(seu) <- c("SCT")
```


```{r}
seu
```
```{r}
DimPlot(seu, label = T, pt.size = 1)+NoLegend()
```



refer to
`https://bustools.github.io/BUS_notebooks_R/slingshot.html#introduction`
Unfortunately, slingshot does not natively support ggplot2. So this is a
function that assigns colors to each cell in base R graphics.

```{r}
#' Assign a color to each cell based on some value
#' 
#' @param cell_vars Vector indicating the value of a variable associated with cells.
#' @param pal_fun Palette function that returns a vector of hex colors, whose
#' argument is the length of such a vector.
#' @param ... Extra arguments for pal_fun.
#' @return A vector of hex colors with one entry for each cell.
cell_pal <- function(cell_vars, pal_fun,...) {
  if (is.numeric(cell_vars)) {
    pal <- pal_fun(100, ...)
    return(pal[cut(cell_vars, breaks = 100)])
  } else {
    categories <- sort(unique(cell_vars))
    pal <- setNames(pal_fun(length(categories), ...), categories)
    return(pal[cell_vars])
  }
}
```

```{r}
cell_colors <- cell_pal(seu$celltype, brewer_pal("qual", "Set2"))
```

```{r}
cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )
```

## 1.1 Generate slingshot object using seurat

```{r}
seu.sc <- as.SingleCellExperiment(seu)
table(seu.sc$ident)

```

```{r}
save(seu.sc, #intitial sce object after conversion
     cell_colors, # default color
     cb_palette, ##  color for different assigned celltype
     file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_renamed_sce.Rdata")
```


Resume
```{r}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
library(scCustomize)
library(destiny)
library(plot3D)
library(rgl)
set.seed(12)
theme_set(theme_bw())

load(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_renamed_sce.Rdata")
seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```

```{r}
source("./PseudotimeAnalysis.r")
```

From our previous notebook using scVelo, we identified GE1 as the start, with GE3, GE4 and GE7 as end points.

```{r}
sce  <- slingshot(seu.sc, clusterLabels = seu.sc$ident, reducedDim = 'UMAP',
                  start.clus = c("GE1"), 
                  end.clus = c("GE3", "GE4", "GE7"),
                  approx_points = 150)               
```
You may use our previous color apporoach using `cb_palette`.

```{r}
SlingshotDataSet(sce)
plot(reducedDims(sce)$UMAP,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.2, label = T)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

## 1.2 Probing with Diffusion map

```{r}
library(destiny)

curve(find_dm_k(n), 0, 5624, xname = 'n')
```
```{r}
set.seed(123)
find_sigmas(sce@int_colData@listData[["reducedDims"]]@listData[["HARMONY"]])
```
```{r}

run_diffMap <- function(data=data, condition=condition,sigma="local", k = 20)
  {
  destinyObj <- as.ExpressionSet(as.data.frame(t(data))) 
  destinyObj$condition <- factor(condition) 
  dm <- DiffusionMap(destinyObj, sigma, k) 
  return(dm)

}

plot_eigenVal <- function(dm=dm){ 
  linepad <- .5 
  plot( 
    eigenvalues(dm), 
    ylim = 0:1,
    pch = 20,
    xlab ='Diffusion component (DC)', 
    ylab ='Eigenvalue'
  )
}

plot_dm_3D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, size=0.5, legend = NULL){
  cond <- factor(condition) 
  col <- factor(condition) 
  levels(col) <- colours 
  col <- as.vector(col) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame(
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]] 
    ) 
  colnames(data) <- DCs
  plot3d( 
    data, 
    bg=col, 
    col=col, 
    size=size, 
    box = FALSE 
    )
  if (is.null(legend)==FALSE){ 
    legend3d("topright", legend = unique(condition), pch = 21, pt.bg = unique(col), cex=1.5, inset=c(0.02), bty = "n") 
     }
}

plot_dm_3D_in_2D <- function(dm=dm, dc=c(1:3), condition=condition, colours=colours, outline.color = "grey20", size=1, pch = 21, theta = 40, phi = 40, bty ="b"){
  cond <- factor(condition) 
  cols <- factor(condition) 
  levels(cols) <- colours 
  cols <- as.vector(cols) 
  DCs <- paste("DC",dc, sep="")
  data <- data.frame( 
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]]
    )
  colnames(data) <- DCs
  if(pch == 21){
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              bg = cols,
              col = outline.color, 
              pch=pch,
              cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta =theta, phi = phi, bty = bty)
    } else {
      scatter3D(x=as.matrix(data[,DCs[1]]),
                y=as.matrix(data[,DCs[2]]), 
                z=as.matrix(data[,DCs[3]]),
                colvar =as.numeric(factor(condition)), col = colours,
                pch=pch, cex = size,
                xlab = DCs[1],
                ylab = DCs[2],
                zlab = DCs[3],
                theta =theta,
                phi = phi,
                bty = bty)
    }
}


plot_dm_3D_in_2D_2 <- function(dm, dc=c(1:3), condition, colours, outline.color = "grey20", size=1, pch = 21, theta = 40, phi = 40, bty ="b", alpha=1){
  
  cond <- factor(condition) 
  cols <- factor(condition) 
  levels(cols) <- colours 
  cols <- as.vector(cols) 
  
  # Adjust colors for transparency
  cols <- grDevices::adjustcolor(cols, alpha=alpha)
  outline.color <- grDevices::adjustcolor(outline.color, alpha=alpha)
  colours <- grDevices::adjustcolor(colours, alpha=alpha)
  
  DCs <- paste("DC",dc, sep="")
  data <- data.frame( 
    dm@eigenvectors[,DCs[1]], 
    dm@eigenvectors[,DCs[2]], 
    dm@eigenvectors[,DCs[3]]
  )
  colnames(data) <- DCs
  
  if(pch == 21){
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              bg = cols,
              col = outline.color, 
              pch=pch,
              cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta = theta, phi = phi, bty = bty)
  } else {
    scatter3D(x=as.matrix(data[,DCs[1]]),
              y=as.matrix(data[,DCs[2]]), 
              z=as.matrix(data[,DCs[3]]),
              colvar = as.numeric(factor(condition)), col = colours,
              pch=pch, cex = size,
              xlab = DCs[1],
              ylab = DCs[2],
              zlab = DCs[3],
              theta = theta,
              phi = phi,
              bty = bty)
  }
}



get_pseudotime <- function(pseudotime, wthres=wthres){
    pseudoT <- list()
    for(lineage in 1:length(pseudotime@metadata[["curves"]]))local({
        curve <- pseudotime@metadata[["curves"]][[lineage]]
        lambda <- curve$lambda
        weight <- curve$w
        ps <- curve$lambda
        ps[weight < wthres] <- NA
        ps <- rankKeepNA(ps)
        pseudoT[[lineage]] <<- ps
    })
    df <- t(do.call("rbind",pseudoT))
    colnames(df) <- names(pseudotime@metadata[["curves"]])
    return(df)
}

```


```{r}
sce_dm <- run_diffMap(t(seu@reductions$harmony@cell.embeddings), colData(sce)$ident, sigma = 29.8, k = 1000) 
plot_eigenVal(dm = sce_dm)
```
```{r, fig.height=20, fig.width=20}
my_color<-cb_palette
names(my_color) <- levels(seu)
#plot.DiffusionMap(sce_dm,dims = 1:2,col_by = phenoData(dataset(sce_dm)))
```
```{r}
sce_lineage <- slingshot(sce_dm@eigenvectors[,c(1:4)], clusterLabels =factor(colData(sce)$ident), 
                         start.clus = 'GE1',
                         end.clus = c("GE3", "GE4", "GE7"),
                         allow.breaks=FALSE, maxit = 1000, shrink.method ="density", thresh = 0.001, extend = "n")
```

```{r}
SlingshotDataSet(sce_lineage)
```


```{r, fig.height=20, fig.width=20}
splom(~sce_dm@eigenvectors[, 1:5], groups =sce$ident, col = my_color, main ="sce_lineage",key = list(space="right", 
                                                                                                      points = list(pch = 19, col =my_color), 
                                                                                                      text = list(c(levels(sce$ident)))))

ggsave()
```




```{r}
plot_dm_3D(dm = sce_dm, dc = c(1,2,3), condition = seu@active.ident, colour = my_color, size = 3) 
rglwidget()
rgl.postscript('./OutputFigure/3dplot.pdf', fmt = 'pdf')
```


Get pseudotime with slingshot
```{r}
#Get pseudotime with slingshot
sce_pseudotime <- get_pseudotime(sce_lineage, wthres = 0.90)
rownames(sce_pseudotime) <- colnames(x = sce)

lineage1 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 1]), 1])
lineage2 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 2]), 2])
lineage3 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 3]), 3])
lineage4 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 4]), 4])
lineage5 <- sort(sce_pseudotime[!is.na(sce_pseudotime[, 5]), 5])

```

```{r}
length(lineage1)
```
```{r}
curves <- slingCurves(sce_lineage)

```


## 1.3 Visualization 2
```{r}
source("./PseudotimeAnalysis.r")
```

```{r, fig.height=10, fig.width=10}
plot_dm_3D_in_2D(dm = sce_dm, dc = c(1, 2, 3), condition = seu@active.ident, 
                 colour = my_color, size = 1, theta = 90, phi = 150, pch = 21, bty = "g")
```





Highlight each lineage
```{r}
highlight_cells_per_lineage_dmplot3D_2 <- function(object, dm=dm, dc=c(1:3), highlight.cells, condition=condition, colours=colours, lineage.curve=NULL, size=1, pch = 21, outline.color = "#E5E4E2", bg.color = "#E5E4E2", title="Diffusion map",phi = 40, theta = 40, add.curve=NULL, bty ="b"){
    
    # Adjusting the transparency for bg.color and outline.color
    bg.color <- grDevices::adjustcolor(bg.color, alpha.f = 0.7)
    outline.color <- grDevices::adjustcolor(outline.color, alpha.f = 0.2)
    
    cell.use.s <- highlight.cells
    
    cols <- factor(condition)
    levels(cols) <- colours
    cols <- as.vector(cols)
    DCs <- paste("DC",dc, sep="")
    
    data <- data.frame(
    dm@eigenvectors[,DCs[1]],
    dm@eigenvectors[,DCs[2]],
    dm@eigenvectors[,DCs[3]],
    Cell=row.names(object@reductions$pca@cell.embeddings),
    Ident=condition,
    Color=cols
    )
    colnames(data) <- c(DCs,"Cell","ident","Color")
    
    data2 = subset(data, data$Cell %in% cell.use.s)
    
    scatter3D(x=as.matrix(data[,DCs[1]]), 
    y=as.matrix(data[,DCs[2]]), 
    z=as.matrix(data[,DCs[3]]), 
    bg = bg.color, 
    col = outline.color, 
    pch=pch, 
    cex = size, 
    xlab = DCs[1], 
    ylab = DCs[2], 
    zlab = DCs[3], 
    theta =theta, 
    phi = phi, main =title, bty = bty)
    
    scatter3D(x=as.matrix(data2[,DCs[1]]), y=as.matrix(data2[,DCs[2]]), z=as.matrix(data2[,DCs[3]]), bg = as.matrix(data2[,"Color"]), 
    col = outline.color, pch=pch, cex = size, xlab = DCs[1], ylab = DCs[2], zlab = DCs[3], theta =theta, phi = phi, add=TRUE, bty = bty)
    
    if(!is.null(add.curve)){
        c <- add.curve
        scatter3D(x = c$s[c$ord,DCs[1]], y =c$s[c$ord,DCs[2]], z = c$s[c$ord,DCs[3]], col = "black", type = "l", ticktype = "detailed", lwd = 4, add = TRUE, xlab = DCs[1], ylab = DCs[2], zlab = DCs[3], theta =theta, phi = phi, bty = bty)
    }
    
}

```



```{r}
glist <- list()
celllist <- list(names(lineage1), names(lineage2), names(lineage3), names(lineage4), names(lineage5))
for (i in 1:5){
	# Start the png device
  filename <- paste0("./OutputFigure/JE_1st_impute_slingshot_phi40_Branch_", i, ".png")
  png(filename, width=1800, height=1800)
  cell.use.s <- celllist[[i]]
  title <- paste("Branch", i, sep="")
  g <- highlight_cells_per_lineage_dmplot3D_2(object = seu, dm = sce_dm, dc = c(1, 2, 3), 
                                            condition = seu@active.ident, colours = my_color,
                                            highlight.cells = cell.use.s, size=0.5,theta = 90,  phi = 10, bty = "n", title = title)
  g <- g + scatter3D(x = curves$Lineage1$s[curves$Lineage1$ord, 1], 
          y = curves$Lineage1$s[curves$Lineage1$ord, 2], 
          z = curves$Lineage1$s[curves$Lineage1$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage2$s[curves$Lineage2$ord, 1], 
          y = curves$Lineage2$s[curves$Lineage2$ord, 2], 
          z = curves$Lineage2$s[curves$Lineage2$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage3$s[curves$Lineage3$ord, 1], 
          y = curves$Lineage3$s[curves$Lineage3$ord, 2], 
          z = curves$Lineage3$s[curves$Lineage3$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage4$s[curves$Lineage4$ord, 1], 
          y = curves$Lineage4$s[curves$Lineage4$ord, 2], 
          z = curves$Lineage4$s[curves$Lineage4$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)
  g <- g + scatter3D(x = curves$Lineage5$s[curves$Lineage5$ord, 1], 
          y = curves$Lineage5$s[curves$Lineage5$ord, 2], 
          z = curves$Lineage5$s[curves$Lineage5$ord, 3], 
          col = "black", type = "l", ticktype = "detailed", lwd = 1, add = TRUE)

  glist[[i]] <- g
    # Close the graphics device, saving the plot
  dev.off()
}
```

```{r}

highlight_cells_per_lineage_UMAPplot <- function(object, highlight.cells, pt.size = 2, pt.shape = 21, outline.color = "grey30", bg.color = "#d3d3d3", hi.color, outline.size = 0.2,title="Umap"){
    
    cell.use.s <- highlight.cells
    
    data.plot=data.frame(object@reductions$umap@cell.embeddings,row.names(object@reductions$pca@cell.embeddings),object@active.ident)
    colnames(data.plot) <- c("UMAP_1","UMAP_2","Cell","ident")
    
    g=ggplot(data.plot,aes(x=UMAP_1,y=UMAP_2))
    g=g+geom_point(size=pt.size, shape=pt.shape, color=outline.color, fill=bg.color, stroke = outline.size)
    g=g+geom_point(data=subset(data.plot, data.plot$Cell %in% cell.use.s), aes(fill =ident), color=outline.color, size=pt.size, shape=pt.shape, stroke = outline.size)
    g=g+scale_fill_manual(values = hi.color,name="")+ggtitle(title)
    g=g+theme_grey()+
    theme(
    axis.text=element_text(size=16),
    axis.title=element_text(size=16),
    legend.text = element_text(size =16),
    legend.title = element_text(size =16 ,face="bold"),
    plot.title = element_text(size=18, face="bold", hjust = 0.5),
    aspect.ratio=1
    )
    return(g)
}

```



```{r}
glist <- list()
celllist <- list(names(lineage1), names(lineage2),names(lineage3), names(lineage4),names(lineage5))
for (i in 1:5){
  cell.use.s <- celllist[[i]]
  title <- paste("Branch", i, sep="")
  g <- highlight_cells_per_lineage_UMAPplot(seu, highlight.cells = cell.use.s, hi.color = my_color, title = title)
  glist[[i]] <- g
}
```
```{r, fig.width=20, fig.height=20}
plot_grid(glist[[1]], glist[[2]], glist[[3]], glist[[4]], glist[[5]], ncol = 3)
ggsave(filename = "./OutputFigure/pJE_1st_impute_slingshot_lineage_umap.png", width = 30, height = 30)
```






## Old codes ###
## 1.3 Initial plot for all Lineages

By default, you can use cell_colors for celltype color options

```{r}
plot(reducedDims(sce)$UMAP,  col = cell_colors, pch=16, asp = 1, cex=0.5)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

```{r}
plot(reducedDims(sce)$PCA,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.5)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

You may also use our previous color apporoach using `cb_palette`.

```{r}
plot(reducedDims(sce)$UMAP,  col =cb_palette[sce$celltype], pch=16, asp = 1, cex=0.2)                       
lines(SlingshotDataSet(sce), lwd=2, type = 'lineage', col='black')
```

```{r}
plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2)                         
lines(SlingshotDataSet(sce), lwd=2,  col='black')
```

You may set alpha (from scales package). With bigger `cex` and smaller
`alpha`.

```{r}
plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette[sce$celltype],0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)                         
lines(SlingshotDataSet(sce), lwd=2,  col='black')
```

Obviously, we will use UMAP for the downstream plot Below, we will
present how to "hide" lineage you don't want to show.

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = i)
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}



```

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))
for (i in nms) {
  colors <- pal[cut(pt[,i], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = i)
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black')
}



```

When running `SlingshotDataSet(sce)` we will see the composite of
different lineages: lineages: 4 Lineage1: GE1 GE8 GE5 GE9 GE7\
Lineage2: GE1 GE8 GE2 GE6\
Lineage3: GE1 GE8 GE4\
Lineage4: GE1 GE8 GE3

## 1.4 Pseudotime plot for Lienage 1

For instance, if you are especially interested in GE7 related, ie,
Lineage 1, you may run the following commands for plot

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage1"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage1")
lines(SlingshotDataSet(sce), linInd = 1)



```

```{r}

pt <- slingPseudotime(sce)

nms <- colnames(pt)

pal <- viridis(100, end = 0.95)
par(mfrow = c(1, 2))

colors <- pal[cut(pt[,"Lineage1"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage1", ylim=c(-10,10))
lines(SlingshotDataSet(sce), linInd = 1)



plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2,ylim=c(-10,10))  
lines(SlingshotDataSet(sce), linInd = 1, type="l")
```

From the above plot, we can clearly see the path for Lineage 1: GE1 GE8
GE5 GE9 GE7\
If you don't want to see the color for other unwanted cluster, you may
try this "dumb" approach, by coloring other cluster as gray (#808080)

```{r}
cb_palette_Lineage1 <- c("#F8766D", #GE1
                "#C0C0C0", #GE2
                "#C0C0C0", #GE3
                "#C0C0C0", #GE4
                "#00C19F", #GE5
                "#C0C0C0", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3"  #GE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage1[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 1, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage1[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 1)
```

## 1.5 Pseudotime plot for Lienage 4

For instance, if you are especially interested in GE3 related, ie,
Lineage 4, you may run the following commands for plot

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage4"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage4")
lines(SlingshotDataSet(sce), 
      linInd = 4 #change to 4 if you are working with Lineage4
      )




```

```{r}
nc <- 3
pt <- slingPseudotime(sce)

nms <- colnames(pt)
nr <- ceiling(length(nms)/nc)
pal <- viridis(100, end = 0.95)
par(mfrow = c(nr, nc))

colors <- pal[cut(pt[,"Lineage4"], breaks = 100)]
plot(reducedDims(sce)$UMAP, col = colors, pch = 16, cex = 0.5, main = "Lineage4")
lines(SlingshotDataSet(sce), linInd = 4)


plot(reducedDims(sce)$UMAP,  
     col =cb_palette[sce$celltype],
     pch=16, 
     asp = 1, 
     cex=0.2)  
lines(SlingshotDataSet(sce), linInd = 4, type="l")
```

From the above plot, we can clearly see the path for Lineage4: GE1 GE8
GE3\
If you don't want to see the color for other unwanted cluster, you may
try this "dumb" approach, by coloring other cluster as gray (#808080)

```{r}

cb_palette_Lineage4 <- c("#F8766D", #JE1
                "#C0C0C0", #JE2
                "#93AA00", #JE3
                "#C0C0C0", #JE4
                "#C0C0C0", #JE5
                "#C0C0C0", #JE6
                "#C0C0C0", #JE7
                "#DB72FB", #JE8
                "#C0C0C0" #JE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage4[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 4, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage4[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 4)
```

Lineage2: GE1 GE8 GE2 GE6

```{r}

cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )

cb_palette_Lineage2 <- c("#F8766D", #JE1
                "#D39200", #JE2
                "#C0C0C0", #JE3
                "#C0C0C0", #JE4
                "#C0C0C0", #JE5
                "#00B9E3", #JE6
                "#C0C0C0", #JE7
                "#DB72FB", #JE8
                "#C0C0C0" #JE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage2[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 2, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage2[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 2)
```

Lineage3: GE1 GE8 GE4

```{r}

cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3
                "#00BA38", #GE4
                "#00C19F", #GE5
                "#00B9E3", #GE6
                "#619CFF", #GE7
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )

cb_palette_Lineage3 <- c("#F8766D", #JE1
                "#C0C0C0", #JE2
                "#C0C0C0", #JE3
                "#00BA38", #JE4
                "#C0C0C0", #JE5
                "#C0C0C0", #JE6
                "#C0C0C0", #JE7
                "#DB72FB", #JE8
                "#C0C0C0" #JE9
                )

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage3[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 3, type="l")

plot(reducedDims(sce)$UMAP,  
     col = alpha(cb_palette_Lineage3[sce$celltype], 0.5),
     pch=16, 
     asp = 1, 
     cex=0.5)  
lines(SlingshotDataSet(sce), linInd = 3)
```

## 1.5 Pseudotime analysis

```{r}
saveRDS(sce, file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot.rds")

```

#2 DE
Let's look at which genes are differentially expressed along one
of those lineages (linage 1, towarding Odam+ population). In dynverse,
feature (gene) importance is calculated by using gene expression to
predict pseudotime value with random forest and finding genes that
contribute the most to the accuracy of the response. Since it's really
not straightforward to convert existing pseudotime results to dynverse
format, it would be easier to build a random forest model. Variable
importance will be calculated for the top 300 highly variable genes
here, with tidymodels. This is to make the code run faster. There are
other methods of trajectory DE as well, which may be more appropriate
but more time consuming to run, such as tradeSeq and SpatialDE (when run
with one dimension). \## 2.1 Tidymodel approach Load necessary packages
and datasets

```{r}
gc()
rm(list=ls())
library(rsample)
library(tidymodels)
library(RColorBrewer)
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
set.seed(12)
theme_set(theme_bw())

seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
sce <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot.rds")
```

```{r}
# Get top highly variable genes
seu <- SCTransform(seu, verbose = FALSE)
top_hvg <- HVFInfo(object = seu, assay = "SCT") %>% 
  mutate(., bc = rownames(.)) %>% 
  arrange(desc(residual_variance)) %>% 
  top_n(300, residual_variance) %>% 
  pull(bc)
# Prepare data for random forest
dat_use <- t(GetAssayData(seu, slot = "data")[top_hvg,])
```

## 2.2 Driver genes in Lineage 1

```{r}
dat_use_df <- cbind(slingPseudotime(sce)[,1], dat_use) # Do lineage 1, so 1nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])
```

The subset of data is randomly split into training and validation; the
model fitted on the training set will be evaluated on the validation
set.

```{r}

dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)
```

tidymodels is a unified interface to different machine learning models,
a "tidier" version of caret. The code chunk below can easily be adjusted
to use other random forest packages as the back end, so no need to learn
new syntax for those packages.

```{r}

model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train)
```

```{r}
summary(dat_use_df$pseudotime)
```

Now it's time to plot some genes deemed the most important to predicting
pseudotime:

```{r}
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:10]
top_genes_cluster1 <- top_genes
top_genes_cluster1
# Convert to gene symbol
#top_gene_name <- gns$gene_name[match(top_genes, gns$gene)]
```

```{r}
par(mfrow = c(2, 3))
for (i in seq_along(top_genes)) {
  colors <- pal[cut(dat_use[,top_genes[i]], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, 
       pch = 16, cex = 0.5, main = top_genes[i])
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}

```

## 2.2 Driver genes in other lineages

### 2.2.1 Lineage 4

```{r}
dat_use_df <- cbind(slingPseudotime(sce)[,4], dat_use) # Do lineage 4, so 1nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])
```

The subset of data is randomly split into training and validation; the
model fitted on the training set will be evaluated on the validation
set.

```{r}

dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)
```

tidymodels is a unified interface to different machine learning models,
a "tidier" version of caret. The code chunk below can easily be adjusted
to use other random forest packages as the back end, so no need to learn
new syntax for those packages.

```{r}

model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train)
```

```{r}
summary(dat_use_df$pseudotime)
```

Now it's time to plot some genes deemed the most important to predicting
pseudotime:

```{r}
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:10]
top_genes_cluster4 <- top_genes
top_genes_cluster4
# Convert to gene symbol
#top_gene_name <- gns$gene_name[match(top_genes, gns$gene)]
```

```{r}
par(mfrow = c(2, 3))
for (i in seq_along(top_genes)) {
  colors <- pal[cut(dat_use[,top_genes[i]], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, 
       pch = 16, cex = 0.5, main = top_genes[i])
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}

```

### 2.2.2 Lineage 3

```{r}
dat_use_df <- cbind(slingPseudotime(sce)[,3], dat_use) # Do lineage 4, so 1nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])
```

The subset of data is randomly split into training and validation; the
model fitted on the training set will be evaluated on the validation
set.

```{r}

dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)
```

tidymodels is a unified interface to different machine learning models,
a "tidier" version of caret. The code chunk below can easily be adjusted
to use other random forest packages as the back end, so no need to learn
new syntax for those packages.

```{r}

model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train)
```

```{r}
summary(dat_use_df$pseudotime)
```

Now it's time to plot some genes deemed the most important to predicting
pseudotime:

```{r}
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:10]
top_genes_cluster3 <- top_genes
top_genes_cluster3
# Convert to gene symbol
#top_gene_name <- gns$gene_name[match(top_genes, gns$gene)]
```

```{r}
par(mfrow = c(2, 3))
for (i in seq_along(top_genes)) {
  colors <- pal[cut(dat_use[,top_genes[i]], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, 
       pch = 16, cex = 0.5, main = top_genes[i])
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}

```

### 2.2.3 Lineag 2

```{r}
dat_use_df <- cbind(slingPseudotime(sce)[,2], dat_use) # Do lineage 4, so 1nd columnn
colnames(dat_use_df)[1] <- "pseudotime"
dat_use_df <- as.data.frame(dat_use_df[!is.na(dat_use_df[,1]),])
```

The subset of data is randomly split into training and validation; the
model fitted on the training set will be evaluated on the validation
set.

```{r}

dat_split <- initial_split(dat_use_df)
dat_train <- training(dat_split)
dat_val <- testing(dat_split)
```

tidymodels is a unified interface to different machine learning models,
a "tidier" version of caret. The code chunk below can easily be adjusted
to use other random forest packages as the back end, so no need to learn
new syntax for those packages.

```{r}

model <- rand_forest(mtry = 200, trees = 1400, min_n = 15, mode = "regression") %>%
  set_engine("ranger", importance = "impurity", num.threads = 3) %>%
  fit(pseudotime ~ ., data = dat_train)
```

```{r}
summary(dat_use_df$pseudotime)
```

Now it's time to plot some genes deemed the most important to predicting
pseudotime:

```{r}
var_imp <- sort(model$fit$variable.importance, decreasing = TRUE)
top_genes <- names(var_imp)[1:10]
top_genes_cluster2 <- top_genes
top_genes_cluster2
# Convert to gene symbol
#top_gene_name <- gns$gene_name[match(top_genes, gns$gene)]
```

```{r}
par(mfrow = c(2, 3))
for (i in seq_along(top_genes)) {
  colors <- pal[cut(dat_use[,top_genes[i]], breaks = 100)]
  plot(reducedDims(sce)$UMAP, col = colors, 
       pch = 16, cex = 0.5, main = top_genes[i])
  lines(SlingshotDataSet(sce), lwd = 2, col = 'black', type = 'lineages')
}

```

### 2.2.4 Summary Lineage specific driver genes

Generate a dataframe recording the top genes in each lineage

```{r}
top_genes_cluster <- data.frame(
  lineage_1 = unlist(top_genes_cluster1),
  lineage_2 = unlist(top_genes_cluster2),
  lineage_3 = unlist(top_genes_cluster3),
  lineage_4 = unlist(top_genes_cluster4)
)
write.csv(top_genes_cluster, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_genes_20220822.csv", row.names = FALSE)
```

###? Heatmap for these genes

## 2.3 tradeSeq approach

Several threads need to be read before starting this approach:
`https://github.com/statOmics/tradeSeq/issues/56` and
`https://github.com/statOmics/tradeSeq/issues/177`
`https://github.com/statOmics/tradeSeq/issues/111` Also, you need to
read
`https://kstreet13.github.io/bioc2020trajectories/articles/workshopTrajectories.html`
and
`https://statomics.github.io/tradeSeq/articles/fitGAM.html#parallel-computing-1`
\## 2.3.0 Load data and packages

```{r load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(slingshot)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(gridExtra)
library(viridis)
library(readr)
library(Hmisc)
library(ggpubr)
library(rsample)
library(tidymodels)
library(scales)
library(tradeSeq)
library(pheatmap)
set.seed(12)
theme_set(theme_bw())

seu <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
sce <-readRDS(file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot.rds")
```

## 2.3.1 Prepare data

```{r}
crv <- SlingshotDataSet(sce)
countMatrix <- sce@assays@data@listData[["counts"]]
counts <- as.matrix(countMatrix)
rm(countMatrix)
```

## 2.3.2 Fit negative binomial model

```{r}
set.seed(5)
icMat <- evaluateK(counts = counts, 
                   sds = crv, 
                   k = 3:10, 
                   nGenes = 800, 
                   verbose = T)
```

The output graphics are organized into four panels. The left panel plots
a boxplot for each number of knots we wanted to evaluate. The plotted
values are the deviation from a gene's AIC at that specific knot value
from the average AIC of that gene across all the knots under evaluation.
Typically, AIC values are somewhat higher for low number of knots, and
we expect them to decrease as the number of knots gets higher. The two
middle panels plot the average drop in AIC across all genes. The middle
left panel simply plots the average AIC, while the middle right panel
plots the change in AIC relative to the average AIC at the lowest knot
number (here, this is 3 knots, as can also be seen from the plot since
the relative AIC equals 1). Finally, the right panel only plots a subset
of genes where the AIC value changes significantly across the evaluated
number of knots. Here, a significant change is defined as a change in
absolute value of at least 2, but this can be tuned using the aicDiff
argument to evaluateK. For the subset of genes, a barplot is displayed
that shows the number of genes that have their lowest AIC at a specific
knot value.

The middle panels show that the drop in AIC levels off if the number of
knots is increased beyond 8, and we will choose that number of knots to
fit the tradeSeq models.

We then fit the models by running the fitGAM function. By default, the
gene-wise NB-GAM estimates one smoother for every lineage using the
negative binomial distribution. Please refer to the fitGAM vignette
Additional to add additional covariates to the model, speed up
computation or allow for custom normalization, amongst others.

```{r}

pseudotime <- slingPseudotime(crv, na = FALSE)
cellWeights <- slingCurveWeights(crv)
```

Since fitGAM will take quite a lot of time, we utilize paralle model.

```{r}
BPPARAM <- BiocParallel::bpparam()
BPPARAM # lists current options
```

We will then set number of cores to be used.

```{r}
BPPARAM$workers <- 20 # use 20 cores
```

```{r}
sce.trade <- fitGAM(counts = counts, pseudotime = pseudotime, cellWeights = cellWeights,
                 nknots = 8, 
                 verbose = TRUE, 
                 parallel=TRUE, 
                 BPPARAM = BPPARAM)
```

The model may be hard to fit for some genes, and hence the fitting
procedure may not converge for all of the genes in a dataset, especially
in datasets with complex patterns and/or many lineages. You can check
the convergence of each gene as shown below, where a TRUE value
corresponds to a converged model fit, and a FALSE value corresponds to a
model that hasn't been able to converge fully.

```{r}
table(rowData(sce.trade)$tradeSeq$converged)
```

```{r}
saveRDS(sce.trade,  file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq.rds")
```

## 2.3.2 Within-lineage comparisons

## 2.3.2.1 Association of gene expression with pseudotime

A first exploration of the data analysis may consist of checking whether
gene expression is associated with a particular lineage. The statistical
test performed here, implemented in the associationTest function, is
testing the null hypothesis that all smoother coefficients are equal to
each other. This can be interpreted as testing whether the average gene
expression is significantly changing along pseudotime.

```{r}
assoRes <- associationTest(sce.trade)
head(assoRes)
```

## 2.3.2.2 Discovering progenitor marker genes

In order to discover marker genes of the progenitor or differentiated
cell population, researchers may be interested in assessing differential
expression between the progenitor cell population (i.e., the starting
point of a lineage) with the differentiated cell type population (i.e.,
the end point of a lineage). The function startVsEndTest uses a Wald
test to assess the null hypothesis that the average expression at the
starting point of the smoother (progenitor population) is equal to the
average expression at the end point of the smoother (differentiated
population). The test basically involves a comparison between two
smoother coefficients for every lineage. The function startVsEndTest
performs a global test across all lineages by default (i.e. it compares
the start and end positions for all lineages simultaneously), but you
can also assess all lineages separately by setting lineages=TRUE.

First, we want to compare the start end with all lineages
simultaneously.

```{r}
startRes_all <- startVsEndTest(sce.trade)
```

```{r}
head(startRes_all)
startRes_all <- startRes_all[order(startRes_all$pvalue, startRes_all$df),]
head(startRes_all)
```

Let's plot the top10 most dynamic genes

```{r}
for (i in rownames(startRes_all)[1:10]){
  print(i)
  print(
    plotSmoothers(sce.trade, counts, gene = i)
  )
  print(
      FeaturePlot(seu, 
                  features = i,
                  cols = c("grey85", brewer.pal(9,"YlOrRd")),
                  ncol=2, 
                  pt.size = 0.1, label=T)
    )
}


```

Secondly, I want to assess all lineages separately

```{r}
startRes <- startVsEndTest(sce.trade,
                           lineages=TRUE)
```

We then want to check the structure of startRes

```{r}
head(startRes)
```

With the structure of the dataframe startRes, we can access the top10
genes in each lineage

```{r}
topGenes <- list()

for (i in 1:4) {
    # Order the genes based on significance for each lineage
    oStart <- order(startRes[[paste0("waldStat_lineage", i)]], decreasing = TRUE)
    
    # Extract the top 10 genes' names for each lineage
    topGenes[[paste0("lineage", i)]] <- names(sce.trade)[oStart[1:10]]
}

# Convert the list to a dataframe
topGenesDF <- data.frame(
    lineage1 = topGenes$lineage1,
    lineage2 = topGenes$lineage2,
    lineage3 = topGenes$lineage3,
    lineage4 = topGenes$lineage4
)

```

```{r}
write.csv(topGenesDF, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_genes_in_each_lineage_20220822.csv", row.names = FALSE )
```

Plot all topDF genes

```{r}
# Convert the dataframe to a vector of genes
all_genes <- as.vector(t(topGenesDF))

# Loop through each gene and create the desired plots
for (i in all_genes) {
  
  print(i)
  
  # First plot using plotSmoothers
  print(
    plotSmoothers(sce.trade, counts, gene = i)
  )
  
  # Second plot using FeaturePlot
  print(
    FeaturePlot(seu, 
                features = i,
                cols = c("grey85", brewer.pal(9,"YlOrRd")),
                ncol=2, 
                pt.size = 0.1, label=T)
  )
}

```

We can visualize the estimated smoothers for the most significant gene
in lineage 1.

```{r}
oStart <- order(startRes$waldStat_lineage1, #If you care about lineage 1, use waldStat_lineage1
                decreasing = TRUE) # a list returning the most likely genes associated with progenitor along the lineage
sigGeneStart <- names(sce.trade)[oStart[1]] # the most significant we are using `1`
print(sigGeneStart)
plotSmoothers(sce.trade, counts, gene = sigGeneStart)
```

```{r}
plotGeneCount(crv, counts, gene = sigGeneStart, title=sigGeneStart)
```

We can visualize the estimated smoothers for the third most significant
gene in lineage 3.

```{r}

sigGeneStart <- names(sce.trade)[oStart[3]] #  most significant we are using `3`
plotSmoothers(sce.trade, counts, gene = sigGeneStart)
```

```{r}
plotGeneCount(crv, counts, gene = sigGeneStart, title=sigGeneStart)
```

If you want to know the top 10 progenitor genes in lineage 1:

```{r}
names(sce.trade)[oStart[1:10]] 
```

If you want to know "Igfbp7"

```{r}
sigGeneStart <- "Igfbp7" #  most significant we are using `3`
plotSmoothers(sce.trade, counts, gene = sigGeneStart)
```

```{r}
plotGeneCount(crv, counts, gene = sigGeneStart, title=sigGeneStart)
```

If you want ot plot for each lineage, refer to
`https://github.com/statOmics/tradeSeq/issues/167`

## 2.3.2.3 Comparing specific pseudotime values within a lineage

The startVsEndTest compares two points on a lineage, and by default it
is comparing the start point with the end point. However, this is a
specific form of a more general capability of the startVsEndTest to
compare any two points on any lineage. If the interest lies in comparing
any two custom pseudotime values, one can specify this using the
pseudotimeValues arguments in startVsEndTest. For example, below we'd
like to compare the expression for each gene at pseudotime values of 0.8
and 0.1.

```{r}
customRes <- startVsEndTest(sce.trade, pseudotimeValues = c(0.1, 0.8))
```

```{r}
head(customRes)
```

```{r}
test <- customRes[order(-customRes$waldStat),]
test[1:10,]
```

```{r}
plotGeneCount(crv, counts, gene = "Lars2")
plotSmoothers(sce.trade, counts, gene = "Lars2")
```

## 2.3.3 Between-lineage comparisons

## 2.3.3.1 Discovering differentiated cell type markers between lineages

tradeSeq can discover marker genes for the differentiated cell types by
comparing the average expression between end points of the
lineage-specific smoothers. This is implemented in the diffEndTest
function. By default, diffEndTest performs a global test, testing the
null hypothesis that the average expression at the endpoints is equal
for all lineages using a multivariate Wald test. If more than two
lineages are present, one can assess all pairwise comparisons using the
pairwise=TRUE argument. Here we use "End" as points for differentiated
stages.

```{r}
endRes <- diffEndTest(sce.trade, pairwise=TRUE)
```

```{r}
endRes_test <- diffEndTest(sce.trade, pairwise=TRUE)
rm(endRes_test)
```

Before going for the top changed genes, you need to check the headers
for `endRes`

```{r}
endRes[1:3, ]
```

Since we performed pairwise comparison, we need to pay attention to
`waldStat_` to speicify the inter-lineage comparison. Below, we first
write a function to get the top10 genes for a given comparison, then we
implemented it for each pairwise comparison

```{r}
# Function to get top 10 genes for a given comparison
getTopGenes <- function(column_name) {
    o <- order(endRes[[column_name]], decreasing = TRUE)
    return(names(sce.trade)[o[1:10]])
}

# Apply the function for each pairwise comparison
topGenes_1vs2 <- getTopGenes("waldStat_1vs2")
topGenes_1vs3 <- getTopGenes("waldStat_1vs3")
topGenes_1vs4 <- getTopGenes("waldStat_1vs4")
topGenes_2vs3 <- getTopGenes("waldStat_2vs3")
topGenes_2vs4 <- getTopGenes("waldStat_2vs4")
topGenes_3vs4 <- getTopGenes("waldStat_3vs4")

# Build the resulting dataframe
topGenesDF <- data.frame(
    Lineage_1_vs_2 = topGenes_1vs2,
    Lineage_1_vs_3 = topGenes_1vs3,
    Lineage_1_vs_4 = topGenes_1vs4,
    Lineage_2_vs_3 = topGenes_2vs3,
    Lineage_2_vs_4 = topGenes_2vs4,
    Lineage_3_vs_4 = topGenes_3vs4
)

```

```{r}
head(topGenesDF)
```

```{r}
write.csv(topGenesDF, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_for_differentiated_between_lineage_20220822.csv", row.names = FALSE )
```

Plot all topDF genes between lineages

```{r}
# Convert the dataframe to a vector of genes
all_genes <- as.vector(t(topGenesDF))

# Loop through each gene and create the desired plots
for (i in all_genes) {
  
  print(i)
  
  # First plot using plotSmoothers
  print(
    plotSmoothers(sce.trade, counts, gene = i)
  )
  
  # Second plot using FeaturePlot
  print(
    FeaturePlot(seu, 
                features = i,
                cols = c("grey85", brewer.pal(9,"YlOrRd")),
                ncol=2, 
                pt.size = 0.1, label=T)
  )
}

```

Let's check the differential genes along each lineage

```{r}
# Convert the dataframe to a vector of genes
all_genes <- as.vector(t(topGenesDF))

# Loop through each gene and create the desired plots
for (i in all_genes) {
  
  print(i)
  
  # First plot using plotSmoothers
  print(
    plotSmoothers(sce.trade, counts, gene = i)
  )
  
  # Second plot using FeaturePlot
  print(
    plotGeneCount(crv, counts, gene = i, title=i)
  )
}




```

Also,I'll loop through the columns of the topGenesDF dataframe and
adjust the curvesCols according to the comparison. For this, we'll
create a helper function to determine the appropriate color codes for
each pairwise comparison: We will save the outputs in png formats.

```{r}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

# Loop through the columns of topGenesDF and plot each gene
for (comparison in colnames(topGenesDF)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in topGenesDF[[comparison]]) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}

```

Below, we can check the genes in each comparison We can plot the most
significant gene using the plotSmoothers function. Here, we compare
Lineage 1 vs Lineage 3

```{r}
o <- order(endRes$waldStat_1vs3, decreasing = TRUE)
sigGene <- names(sce.trade)[o[1]]

```

```{r}
sigGene
```

```{r}
plotSmoothers(sce.trade, counts, sigGene)
```

```{r}
 names(sce.trade)[o[1:10]]
```

```{r}
plotSmoothers(sce.trade, counts, "Fam57a")
```

OK, if you want to plot gene changes specifically for lineage 1 and 3,
assign the cluster with different color (destination cluster of each
lineage, ie GE7 for lineage 1, GE3 for lineage 3) and transparent other
lineages as below: refer to
`https://github.com/statOmics/tradeSeq/issues/167`

```{r}
cb_palette <- c("#F8766D", #GE1
                "#D39200", #GE2
                "#93AA00", #GE3/Lineage4
                "#00BA38", #GE4/Lineage3
                "#00C19F", #GE5
                "#00B9E3", #GE6/Lineage6
                "#619CFF", #GE7/Lineage1
                "#DB72FB", #GE8
                "#FF61C3" #GE9
                )
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Fam57a", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
ggsave("./OutputFigure/Plot_Fam57a_lineage1vs3.png",  width = 10, height = 5)
```

Of course, you may use the above script to plot any gens you want. For
instance, Krt14 and Krt5

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Krt14", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
ggsave("./OutputFigure/Plot_Krt14_lineage1vs3.png",  width = 10, height = 5)
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Krt5", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
ggsave("./OutputFigure/Plot_Krt5_lineage1vs3.png",  width = 10, height = 5)
```

Krt16

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Krt16", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
ggsave("./OutputFigure/Plot_Krt16_lineage1vs3.png",  width = 10, height = 5)
```

```{r}
curvesCols <- c("#619CFF", "#00B9E3","#00BA38","transparent")
plotSmoothers(sce.trade, counts, "Itgb4", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
ggsave("./OutputFigure/Plot_Itgb4_lineage1vs3.png",  width = 10, height = 5)
```

```{r, two-color plot for different krt genes,fig.height=5, fig.width=15}
FeaturePlot(object = seu, features = c("Fam57a", "Lor"), cols= c("#ddf4f5", "orange", "blue", "green"), blend  = TRUE)

```

```{r, two-color plot for different krt genes,fig.height=5, fig.width=15}
FeaturePlot(object = seu, features = c("Fam57a", "Odam"), cols= c("#ddf4f5", "orange", "blue", "green"), blend  = TRUE)

```

```{r, two-color plot for different krt genes,fig.height=5, fig.width=15}
FeaturePlot(object = seu, features = c("Ptma", "Lor"), cols= c("#ddf4f5", "orange", "blue", "green"), blend  = TRUE)

```

```{r, two-color plot for different krt genes,fig.height=5, fig.width=15}
FeaturePlot(object = seu, features = c("Ptma", "Odam"), cols= c("#ddf4f5", "orange", "blue", "green"), blend  = TRUE)

```

## 2.3.3.2 Discovering genes with different expression patterns

Asides from testing at the level of the differentiated cell type,
researchers may be interested in assessing the expression pattern of a
gene over pseudotime. The function patternTest implements a statistical
method that checks whether the smoothed gene expression is equal along
pseudotime between two or multiple lineages. In practice, we use 100
points, equally distributed along pseudotime, that are compared between
two (or multiple) lineages, and this number can be changed using the
nPoints argument.

```{r}
patternRes <- patternTest(sce.trade, pairwise=TRUE)
```

Confronted with the Error "Error in t(L) %\*% Sigma : non-conformable
arguments", I referred to
`https://github.com/statOmics/tradeSeq/issues/56` and picked the most
abnormal rows containing the fewest non-null values.

```{r}
# Assuming your data is in tradeSeq_data
tradeSeq_data <- rowData(sce.trade)$tradeSeq

# 1. Drop the 'Sigma' column
tradeSeq_data <- tradeSeq_data %>% dplyr::select(-Sigma)
```

```{r}
# 2. Compute the number of non-missing values per row
num_non_missing <- rowSums(!is.na(tradeSeq_data))

# 3. Identify the rows with the fewest non-missing values
min_non_missing <- min(num_non_missing)
rows_with_fewest_non_missing <- which(num_non_missing == min_non_missing)

# Show the rows with the fewest non-missing values
tradeSeq_data[rows_with_fewest_non_missing, ]

```

Again, we performed the patternTest using the sce.trade object with
abnormal rows excluded.

```{r}
patternRes <- patternTest(sce.trade[-rows_with_fewest_non_missing,], pairwise=TRUE)
```

We then check the headers

```{r}
patternRes[1:3, ]
```

We want to see which genes

```{r}
oPat <- order(patternRes$waldStat, 
              decreasing = TRUE)
head(rownames(patternRes)[oPat])
```

List the top 10 genes:

```{r}
rownames(patternRes)[oPat][1:10]
top10_patternRes <- rownames(patternRes)[oPat][1:10]
write.csv(top10_patternRes, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_genes_with_different_pattern_among_all_lineages_20220822.csv", row.names = FALSE )

```

We then check the top10 differential genes among the four lineages

```{r}
curvesCols <- c("#619CFF", "#00B9E3","#00BA38","#93AA00")
    for (gene in top10_patternRes) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_different_pattern_among_all_lineages", "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
```

Below, we first write a function to get the top10 genes for a given
comparison, then we implemented it for each pairwise comparison

```{r}
# Function to get top 10 genes for a given comparison
getTopGenes <- function(column_name) {
    o <- order(patternRes[[column_name]], decreasing = TRUE)
    return(names(sce.trade)[o[1:10]])
}

# Apply the function for each pairwise comparison
topGenes_1vs2 <- getTopGenes("waldStat_1vs2")
topGenes_1vs3 <- getTopGenes("waldStat_1vs3")
topGenes_1vs4 <- getTopGenes("waldStat_1vs4")
topGenes_2vs3 <- getTopGenes("waldStat_2vs3")
topGenes_2vs4 <- getTopGenes("waldStat_2vs4")
topGenes_3vs4 <- getTopGenes("waldStat_3vs4")

# Build the resulting dataframe
topGenesDF_1v1 <- data.frame(
    Lineage_1_vs_2 = topGenes_1vs2,
    Lineage_1_vs_3 = topGenes_1vs3,
    Lineage_1_vs_4 = topGenes_1vs4,
    Lineage_2_vs_3 = topGenes_2vs3,
    Lineage_2_vs_4 = topGenes_2vs4,
    Lineage_3_vs_4 = topGenes_3vs4
)

```

```{r}
head(topGenesDF_1v1)
```

```{r}
write.csv(topGenesDF_1v1, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_genes_with_different_pattern_between_two_lineages_20220822.csv", row.names = FALSE )
```

Also,I'll loop through the columns of the topGenesDF_1v1 dataframe and
adjust the curvesCols according to the comparison. For this, we'll
create a helper function to determine the appropriate color codes for
each pairwise comparison: We will save the outputs in png formats.

```{r}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

# Loop through the columns of topGenesDF_1v1 and plot each gene
for (comparison in colnames(topGenesDF_1v1)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in topGenesDF_1v1[[comparison]]) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_different_pattern_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}

```

## 2.3.3.3 Early drivers of differentiation between lineage

Another question of interest is to find a list of genes that are
differentially expressed between lineages at a particular region,
e.g. around the separation of two or multiple lineages. The function
earlyDETest implements a statistical method to test the null hypothesis
of whether the average gene expression smoothers are equal between
lineages in a region defined by two user-specified knots. Again, the
knots can be visualized with the plotGeneCount function. By selecting
the region covering the first two knot points to test for differential
patterns between the lineages, we check which genes are behaving
differently around the bifurcation point.

```{r}
plotc <- plotGeneCount(curve = crv, counts = counts,
              clusters = apply(slingClusterLabels(crv), 1, which.max),
              models = sce.trade)
plotc
ggsave("./OutputFigure/JE_combined_epi_harmony_SCT_renamed_slingshot_pseudotime_knot.png",plot=plotc, width = 10, height = 8)
```

From the knots in the umap along each lineage, we assumed the
bifurcation happened between knot1 to knot4.

```{r}
earlyDERes <- earlyDETest(sce.trade, knots = c(1, 4),pairwise=TRUE)
```

Still confronted with the error "Error in t(L) %\*% Sigma :
non-conformable arguments"

```{r}
earlyDERes <- earlyDETest(sce.trade[-rows_with_fewest_non_missing,], pairwise = TRUE)
```

We then check the headers

```{r}
earlyDERes[1:3, ]
```

We want to see which genes

```{r}
oPat <- order(earlyDERes$waldStat, 
              decreasing = TRUE)
head(rownames(earlyDERes)[oPat])
```

List the top 10 genes:

```{r}
rownames(earlyDERes)[oPat][1:10]
top10_earlyDERes <- rownames(earlyDERes)[oPat][1:10]
write.csv(top10_earlyDERes, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_EarlyDriver_genes_with_different_pattern_among_all_lineages_20220822.csv", row.names = FALSE )

```

We then plot the top10 early driver genes among the four lineages

```{r}
curvesCols <- c("#619CFF", "#00B9E3","#00BA38","#93AA00")
    for (gene in top10_earlyDERes) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_different_early_driver_among_all_lineages", "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
```

Below, we first write a function to get the top10 genes for a given
comparison, then we implemented it for each pairwise comparison Here we
will pick the top 10 early driver genes differentially activated between
two lineages

```{r}
# Function to get top 10 genes for a given comparison
getTopEDGenes <- function(column_name) {
    o <- order(earlyDERes[[column_name]], decreasing = TRUE)
    return(names(sce.trade)[o[1:10]])
}

# Apply the function for each pairwise comparison
topEDGenes_1vs2 <- getTopGenes("waldStat_1vs2")
topEDGenes_1vs3 <- getTopGenes("waldStat_1vs3")
topEDGenes_1vs4 <- getTopGenes("waldStat_1vs4")
topEDGenes_2vs3 <- getTopGenes("waldStat_2vs3")
topEDGenes_2vs4 <- getTopGenes("waldStat_2vs4")
topEDGenes_3vs4 <- getTopGenes("waldStat_3vs4")

# Build the resulting dataframe
topEDGenes_1v1 <- data.frame(
    Lineage_1_vs_2 = topEDGenes_1vs2,
    Lineage_1_vs_3 = topEDGenes_1vs3,
    Lineage_1_vs_4 = topEDGenes_1vs4,
    Lineage_2_vs_3 = topEDGenes_2vs3,
    Lineage_2_vs_4 = topEDGenes_2vs4,
    Lineage_3_vs_4 = topEDGenes_3vs4
)

```

```{r}
head(topEDGenes_1v1)
```

```{r}
write.csv(topEDGenes_1v1, file = "./OutputTable/JE_combined_epi_harmony_SCT_renamed_slingshot_top_10_early_driver_genes_between_two_lineages_20220822.csv", row.names = FALSE )
```

I also want to get the unique genes in each comparison

```{r}
unique_genes <- lapply(1:ncol(topEDGenes_1v1), function(col_idx) {
  current_genes <- topEDGenes_1v1[[col_idx]]
  other_genes <- unlist(topEDGenes_1v1[,-col_idx])
  unique_genes <- current_genes[!(current_genes %in% other_genes)]
  return(unique_genes)
})

names(unique_genes) <- colnames(topEDGenes_1v1)
```

```{r}
# Find the maximum number of unique genes for all comparisons
max_length <- max(sapply(unique_genes, length))

# Function to fill a vector with NAs up to max_length
fill_NA <- function(vec) {
  c(vec, rep(NA, max_length - length(vec)))
}

# Apply the fill_NA function to each vector in unique_genes
filled_genes <- lapply(unique_genes, fill_NA)

# Convert the list to a dataframe
unique_genes_df <- as.data.frame(filled_genes)

unique_genes_df
```

Also,I'll loop through the columns of the topEDGenes_1v1 dataframe and
adjust the curvesCols according to the comparison. For this, we'll
create a helper function to determine the appropriate color codes for
each pairwise comparison: We will save the outputs in png formats.

```{r}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}
# Loop through the columns of unique_genes_df and plot each gene
for (comparison in colnames(unique_genes_df)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in unique_genes_df[[comparison]]) {
        
        # Omit NA genes
        if (is.na(gene)) next
        
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_unique_early_driver_gene_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}


```

```{r}

library(ggplot2)

# Function to determine color codes for a given pairwise comparison
getCurvesCols <- function(comparison) {
    allCols <- c("#93AA00", "#00BA38", "#00B9E3", "#619CFF")
    transp1 <- "transparent"
    transp2 <- "transparent"
    
    if (comparison == "Lineage_1_vs_2") return(c(allCols[4], allCols[3], transp1, transp2))
    if (comparison == "Lineage_1_vs_3") return(c(allCols[4], transp1, allCols[2], transp2))
    if (comparison == "Lineage_1_vs_4") return(c(allCols[4], transp1, transp2, allCols[1]))
    if (comparison == "Lineage_2_vs_3") return(c(transp1, allCols[3], allCols[2], transp2))
    if (comparison == "Lineage_2_vs_4") return(c(transp1, allCols[3], transp2, allCols[1]))
    if (comparison == "Lineage_3_vs_4") return(c(transp1, transp2, allCols[2], allCols[1]))
}

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

# Loop through the columns of topEDGenes_1v1 and plot each gene
for (comparison in colnames(topEDGenes_1v1)) {
    curvesCols <- getCurvesCols(comparison)
    for (gene in topEDGenes_1v1[[comparison]]) {
        plot_obj <- plotSmoothers(sce.trade, counts, gene, curvesCols = curvesCols, border = FALSE) +
                    scale_color_manual(values = curvesCols)
        
        # Define filename
        filename <- paste(output_dir, "Plot_top10_early_driver_gene_", comparison, "_", gene, ".png", sep = "")
        
        # Save plot to file
        ggsave(filename, plot = plot_obj, width = 10, height = 5)
    }
}

```

Example for specific comparison

Here we want to compare Lineage 1 vs Lineage 3

```{r}
oEarly <- order(earlyDERes$waldStat_1vs3, decreasing = TRUE)
head(rownames(earlyDERes)[oEarly])
```

```{r}
rownames(earlyDERes)[oEarly][1:100] # list the top 100 early driver genes differentially activated between lineage 1 and lineage 3
```

top1

```{r}
rownames(earlyDERes)[oEarly][1]
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = rownames(earlyDERes)[oEarly][1], curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
plotGeneCount(crv, counts, gene = rownames(earlyDERes)[oEarly][1], title=rownames(earlyDERes)[oEarly][1])
```

top2

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = rownames(earlyDERes)[oEarly][2], curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
plotGeneCount(crv, counts, gene = rownames(earlyDERes)[oEarly][2], title=rownames(earlyDERes)[oEarly][2])
```

top3

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = rownames(earlyDERes)[oEarly][3], curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
plotGeneCount(crv, counts, gene = rownames(earlyDERes)[oEarly][3], title=rownames(earlyDERes)[oEarly][3])
```

top4

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = rownames(earlyDERes)[oEarly][4], curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
plotGeneCount(crv, counts, gene = rownames(earlyDERes)[oEarly][4], title=rownames(earlyDERes)[oEarly][4])
```

top5

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = rownames(earlyDERes)[oEarly][5], curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
plotGeneCount(crv, counts, gene = rownames(earlyDERes)[oEarly][5], title=rownames(earlyDERes)[oEarly][5])
```

For a known gene

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = "Il24", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
curvesCols_all <- c("#619CFF", "#00B9E3","#00BA38" , "#93AA00")
plotSmoothers(sce.trade, counts, gene = "Il24", curvesCols = curvesCols_all, border = FALSE) 
plotGeneCount(crv, counts, gene = "Il24", title="Il24")
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = "Gclc", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
curvesCols_all <- c("#619CFF", "#00B9E3","#00BA38" , "#93AA00")
plotSmoothers(sce.trade, counts, gene = "Gclc", curvesCols = curvesCols_all, border = FALSE) 
plotGeneCount(crv, counts, gene = "Gclc", title="Gclc")
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = "Slc7a11", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
curvesCols_all <- c("#619CFF", "#00B9E3","#00BA38" , "#93AA00")
plotSmoothers(sce.trade, counts, gene = "Slc7a11", curvesCols = curvesCols_all, border = FALSE) 
plotGeneCount(crv, counts, gene = "Slc7a11", title="Slc7a11")
```

```{r}
curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
plotSmoothers(sce.trade, counts, gene = "Slc1a5", curvesCols = curvesCols,
              border = FALSE) +
  scale_color_manual(values = curvesCols)
curvesCols_all <- c("#619CFF", "#00B9E3","#00BA38" , "#93AA00")
plotSmoothers(sce.trade, counts, gene = "Slc1a5", curvesCols = curvesCols_all, border = FALSE) 
plotGeneCount(crv, counts, gene = "Slc1a5", title="Slc1a5")
```

OK, let's make a loop for the top10 early driver genes between lineage 1
and lineage 3

```{r}
for (i in rownames(earlyDERes)[oEarly][1:10]){
  i
  curvesCols <- c("#619CFF", "transparent","#00BA38","transparent")
  print(
    plotSmoothers(sce.trade, counts, gene = i, curvesCols = curvesCols,border = FALSE) + scale_color_manual(values = curvesCols)
  )
  print(
    plotGeneCount(crv, counts, gene = i, title=i)
  )
}
```

## 2.3.3.4 Bifurcation heatmap for early driver genes among lineage 1 and 3.

However, we want to know if there are any trends cluster along the
bifurcation.Using hybrid heatmap function, we are trying to see if
slingshot+tradeseq could replace BEAM in monocle2. For instance, we want
to check the trends of the top 200 genes in 100 pseudotime intervals.
Immediately after the last part, the oEarly indicate the early driver
genes differentially activated in lineage 1 or lineage 3. Here we want
to compare Lineage 1 vs Lineage 3. And we will repeat some of the
command lines in the last part.

```{r}
oEarly <- order(earlyDERes$waldStat_1vs3, decreasing = TRUE) # get the most differentially activated genes between lineage 1 and lineage 3
head(rownames(earlyDERes)[oEarly])
```

```{r}
rownames(earlyDERes)[oEarly][1:100] # list the top 100 early driver genes differentially activated between lineage 1 and lineage 3
```

```{r}
yhatSmooth_false <- predictSmooth(sce.trade, gene = rownames(earlyDERes)[oEarly][1:200], #top 200 genes
                                  nPoints = 100, #idealy, I would use 30-100, larger number will make the scale effect not that significant
                                  tidy = FALSE)

```

Since we know that the matrix counts were arranged along the pseudotime,
and we generate the oEarly by the Waldtest between lineage 1 and 3, we
subset and re-arrange the count matrix to generate heatmap using the
hybrid line below:

```{r,fig.height=30, fig.width=20}
heatSmooth <- pheatmap(t(scale(t(yhatSmooth_false[, c(100:1,201:300)]))), # yhatSmooth_false we used nPoints=100, so we choose 1:100 and 201:300 for lineage 1 and 3 respectively
                       cluster_cols = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = TRUE,
                       border_color = FALSE)
ggsave("./OutputFigure/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_1_vs3_bifurcation_heatmap.png", plot=heatSmooth, width = 10, height = 25)
```

If you don't want to cluster rows, rownames for the genes were listed
according to the value of waldTest in a decreasing manner.

```{r,fig.height=30, fig.width=20}
heatSmooth_2 <- pheatmap(t(scale(t(yhatSmooth_false[, c(100:1,201:300)]))), # yhatSmooth_false we used nPoints=100, so we choose 1:100 and 201:300 for lineage 1 and 3 respectively
                       cluster_cols = FALSE,
                       cluster_rows = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = TRUE,
                       border_color = FALSE)
ggsave("./OutputFigure/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_1_vs3_bifurcation_heatmap_unclustered.png", plot=heatSmooth_2, width = 10, height = 25)
```

Here, I'm just not certain about the WaldTest value, so I filter the
genes using the p value.And compare the heatmaps.

First, we regenerate the oEarly order as below: Here we want to compare
Lineage 1 vs Lineage 3

```{r}
earlyDERes_p <- earlyDERes[earlyDERes$pvalue_1vs3<0.001,]
dim(earlyDERes_p)
oEarly_2 <- order(earlyDERes_p$waldStat_1vs3, decreasing = TRUE)
head(rownames(earlyDERes_p)[oEarly_2])
```

```{r}
yhatSmooth_false <- predictSmooth(sce.trade, gene = rownames(earlyDERes_p)[oEarly_2][1:200], nPoints = 100, tidy = FALSE)

```

Since we know that the matrix counts were arranged along the pseudotime,
and we generate the oEarly by the Waldtest between lineage 1 and 3, we
subset and re-arrange the count matrix to generate heatmap using the
hybrid line below:

```{r,fig.height=30, fig.width=20}
heatSmooth <- pheatmap(t(scale(t(yhatSmooth_false[, c(100:1,201:300)]))), # yhatSmooth_false we used nPoints=100, so we choose 1:100 and 201:300 for lineage 1 and 3 respectively
                       cluster_cols = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = TRUE,
                       border_color = FALSE)
```

If you don't want to cluster them,

```{r,fig.height=30, fig.width=20}
heatSmooth_2 <- pheatmap(t(scale(t(yhatSmooth_false[, c(100:1,201:300)]))), # yhatSmooth_false we used nPoints=100, so we choose 1:100 and 201:300 for lineage 1 and 3 respectively
                       cluster_cols = FALSE,
                       cluster_rows = FALSE,
                       show_rownames = TRUE, 
                       show_colnames = TRUE,
                       border_color = FALSE)
```

##2.3.3.5 Loop for Bifurcation heatmap

```{r,fig.height=30, fig.width=20}
# Define the different comparisons based on the naming pattern in earlyDERes
comparisons <- c("1vs3", "1vs4", "1vs2", "2vs3", "2vs4", "3vs4")

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

for (comp in comparisons) {
    # Order by the waldStat column for this comparison
    waldStat_col <- paste0("waldStat_", comp)
    oEarly <- order(earlyDERes[[waldStat_col]], decreasing = TRUE)
    
    # Predict smooth values
    yhatSmooth_false <- predictSmooth(sce.trade, gene = rownames(earlyDERes)[oEarly][1:200], nPoints = 100, tidy = FALSE)
    
    # Calculate the columns to extract based on current lineage comparison
    lineages <- as.numeric(strsplit(comp, "vs")[[1]])
    idx1 <- seq(lineages[1]*100, by = -1, length.out = 100)
    idx2 <- seq((lineages[2]-1)*100 + 1, lineages[2]*100)
    
    # Generate the heatmaps
    heatSmooth <- pheatmap(t(scale(t(yhatSmooth_false[, c(idx1, idx2)]))),
                           cluster_cols = FALSE,
                           show_rownames = TRUE, 
                           show_colnames = TRUE,
                           border_color = FALSE)
    
    # Save the clustered heatmap
    filename <- paste(output_dir, "JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_", comp, "_bifurcation_heatmap.png", sep = "")
    ggsave(filename, plot = heatSmooth, width = 10, height = 25)
    
    # Generate the unclustered heatmap
    heatSmooth_2 <- pheatmap(t(scale(t(yhatSmooth_false[, c(idx1, idx2)]))),
                             cluster_cols = FALSE,
                             cluster_rows = FALSE,
                             show_rownames = TRUE, 
                             show_colnames = TRUE,
                             border_color = FALSE)
    
    # Save the unclustered heatmap
    filename_2 <- paste(output_dir, "JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_", comp, "_bifurcation_heatmap_unclustered.png", sep = "")
    ggsave(filename_2, plot = heatSmooth_2, width = 10, height = 25)
}

```

If considering the p value threshold

```{r,fig.height=30, fig.width=20}
# Define the different comparisons based on the naming pattern in earlyDERes
comparisons <- c("1vs3", "1vs4", "1vs2", "2vs3", "2vs4", "3vs4")

# Output directory
output_dir <- "./OutputFigure/"

# Check if the output directory exists, if not, create it
if (!dir.exists(output_dir)) {
    dir.create(output_dir)
}

for (comp in comparisons) {
    # Filter and order by the waldStat column for this comparison
    waldStat_col <- paste0("waldStat_", comp)
    pvalue_col <- paste0("pvalue_", comp)
    
    earlyDERes_filtered <- earlyDERes[earlyDERes[[pvalue_col]] < 0.001, ]
    
    # Get the number of genes that pass the p-value threshold
    num_genes <- nrow(earlyDERes_filtered)
    
    # Use the minimum of 200 or the number of genes that passed the threshold
    genes_to_use <- min(200, num_genes)
    
    oEarly <- order(earlyDERes_filtered[[waldStat_col]], decreasing = TRUE)
    
    # Predict smooth values
    yhatSmooth_false <- predictSmooth(sce.trade, gene = rownames(earlyDERes_filtered)[oEarly][1:genes_to_use], nPoints = 100, tidy = FALSE)
    
    # Calculate the columns to extract based on current lineage comparison
    lineages <- as.numeric(strsplit(comp, "vs")[[1]])
    idx1 <- seq(lineages[1]*100, by = -1, length.out = 100)
    idx2 <- seq((lineages[2]-1)*100 + 1, lineages[2]*100)
    
    # Generate the heatmaps
    heatSmooth <- pheatmap(t(scale(t(yhatSmooth_false[, c(idx1, idx2)]))),
                           cluster_cols = FALSE,
                           show_rownames = TRUE, 
                           show_colnames = TRUE,
                           border_color = FALSE)
    
    # Save the clustered heatmap
    filename <- paste(output_dir, "JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_", comp, "_bifurcation_heatmap.png", sep = "")
    ggsave(filename, plot = heatSmooth, width = 10, height = 25)
    
    # Generate the unclustered heatmap
    heatSmooth_2 <- pheatmap(t(scale(t(yhatSmooth_false[, c(idx1, idx2)]))),
                             cluster_cols = FALSE,
                             cluster_rows = FALSE,
                             show_rownames = TRUE, 
                             show_colnames = TRUE,
                             border_color = FALSE)
    
    # Save the unclustered heatmap
    filename_2 <- paste(output_dir, "JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq_lineage_", comp, "_bifurcation_heatmap_unclustered.png", sep = "")
    ggsave(filename_2, plot = heatSmooth_2, width = 10, height = 25)
}


```

## 2.x Save Rdata

```{r}
save(sce.trade, #sce fitGAM
     assoRes, # lineage associated genes
     earlyDERes, # lineage specific driver genes between lineages
     startRes,#progenitor marker genes within all lineages
     startRes_all, #progenitor marker genes within each lineages
     customRes, # lineage specific genes within a specific pseudotime point of a specific lineage
     endRes, # differentiated cell type markers between lineages
     patternRes, ##  genes with different expression patterns
     file = "./SeuratObject/JE_combined_epi_harmony_SCT_renamed_slingshot_tradeSeq.Rdata")
```


## 2.y Plot
```{r, marker for cluster krt,fig.height=12, fig.width=10}
library(RColorBrewer)
DefaultAssay(seu) <- "SCT"
FeaturePlot(seu, features = c("Odam","Nedd4",
                                          "Klk10", "Atp1a1", "Sbsn","Fam57a"),
            cols = c("grey85",brewer.pal(9,"YlOrRd")),
            ncol=2, pt.size = 0.1,  label=TRUE)+xlim(-7,7)+ylim(-7,7)
#ggsave("./OutputFigure/palate_epi_MAGIC_periderm_MAGIC_tf_feature_MAGIC.png",  width = 10, height = 24)
```
```{r, marker for cluster krt,fig.height=24, fig.width=15}
DefaultAssay(seu) <- "MAGIC_SCT"
FeaturePlot(seu, features = c("Odam","Nedd4",
                                          "Klk10", "Atp1a1", "Sbsn","Fam57a"),
            cols = c("grey85",brewer.pal(9,"YlOrRd")),
            ncol=2, pt.size = 0.1, split.by = "orig.ident", label=TRUE)+xlim(-7,7)+ylim(-7,7)
#ggsave("./OutputFigure/palate_epi_MAGIC_periderm_MAGIC_pe7_feature_MAGIC.png",  width = 10, height = 24)
```
```{r, marker for cluster krt,fig.height=24, fig.width=15}
DefaultAssay(seu) <- "MAGIC_SCT"
FeaturePlot(seu, features = c("Odam","Nedd4",
                                          "Krtadp", "Krt5", "Krt14","Lor"),
            cols = c("grey85",brewer.pal(9,"YlOrRd")),
            ncol=2, pt.size = 0.1, split.by = "orig.ident", label=TRUE)+xlim(-7,7)+ylim(-7,7)
#ggsave("./OutputFigure/palate_epi_MAGIC_periderm_MAGIC_pe7_feature_MAGIC.png",  width = 10, height = 24)
```

###STOP HERE \###

## 2.3.4 Clustering of genes according to their expression pattern

## 2.3.4.1 Extracting fitted values to use with any clustering method

While tradeSeq provides readily implemented functionality to cluster
genes using the clusterExperiment package, users may want to cluster
genes using other clustering methods. Fitted values from the tradeSeq
models, which can subsequently be used for clustering, can be extracted
using the predictSmooth or predictCells functions. predictCells predicts
the estimated expression for each cell (i.e.it is a fitted value for
that cell) on the count scale. predictSmooth, instead, returns values of
the estimated smoother on a grid of pseudotimes. Either of these may be
used as input to a clustering method; roughly, predictSmooth may be
considered as a cleaner version of predictCells. By default
predictSmooth returns a data.frame specifying the estimated smoother for
each gene. A matrix containing these values, where each row is a gene,
and each column is a point along the lineage on the uniform grid, can be
obtained by setting tidy=FALSE in predictSmooth.

```{r}
yhat <- predictCells(models = sce.trade, gene = "Odam")

ysmooth <- predictSmooth(models = sce.trade, gene = "Odam", nPoints = 40, tidy=FALSE)
```

```{r}
zhat <- predictCells(models = sce.trade, gene = "Itgb4")

zsmooth <- predictSmooth(models = sce.trade, gene = "Itgb4", nPoints = 40)
```

## 2.3.4.2 Clustering using RSEC, clusterExperiment

tradeSeq provides the functionality to cluster genes according to their
expression pattern along the lineages with the clusterExpressionPatterns
function. A number of equally spaced points for every lineage are
selected to perform the clustering, and the number of points can be
selected with the nPoints argument. The genes argument specifies which
genes you want to cluster (e.g., all genes with differential expression
patterns). Here, we use 20 points along each lineage to cluster the all
genes in the dataset. The clustering itself occurs by the
clusterExperiment package (Risso et al. 2018), hence the user may select
any clustering algorithm that's built into that package, or custom
clustering algorithms implemented by the user. For a list of built-in
clustering algorithms within clusterExperiment, run
clusterExperiment::listBuiltInFunctions() on the command line. For more
details about RSEC and clusterExperiment, refer to
`https://bioconductor.org/packages/release/bioc/vignettes/clusterExperiment/inst/doc/clusterExperimentTutorial.html`

```{r}
library(clusterExperiment)
```

```{r}
nPointsClus <- 20
clusPat <- clusterExpressionPatterns(sce.trade, nPoints = nPointsClus,
                                     genes = rownames(counts)[1:1000])
```

```{r}
table(is.na(counts))
```

```{r}
clusterLabels <- primaryCluster(clusPat$rsec)
```

The first 4 clusters can be visualized using the normalized expression
upon which the clustering is based. Please note that the code below
would only works for a trajectory with two lineages. Modify the code
appropriately if using with a dataset with 3 lineages or more.

```{r}
cUniq <- unique(clusterLabels)
cUniq <- cUniq[!cUniq == -1] # remove unclustered genes

for (xx in cUniq[1:4]) {
  cId <- which(clusterLabels == xx)
  p <- ggplot(data = data.frame(x = 1:nPointsClus,
                                y = rep(range(clusPat$yhatScaled[cId, ]),
                                        nPointsClus / 2)),
              aes(x = x, y = y)) +
    geom_point(alpha = 0.5) +
    labs(title = paste0("Cluster ", xx),  x = "Pseudotime", y = "Normalized expression") +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5))
  
  for (ii in 1:length(cId)) {
    geneId <- rownames(clusPat$yhatScaled)[cId[ii]]
    p <- p +
      geom_line(data = data.frame(x = rep(1:nPointsClus, 2),
                                  y = clusPat$yhatScaled[geneId, ],
                                  lineage = rep(0:1, each = nPointsClus)),
                aes(col = as.character(lineage), group = lineage), lwd = 0.8)
  }
  p <- p + guides(color = FALSE) +
    scale_color_manual(values = c("orange", "darkseagreen3"),
                       breaks = c("0", "1"))  
  print(p)
}
```
