---
title: "Mouse JE repair scRNA harmony epi imputation using MAGIC"
author: "Huan"
date: "12/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = TRUE,
  warning = FALSE,
  cache = TRUE
)
```

This note was used for imputation exploration for JE_epi.
We employed the following approaches: MAGIC

# 1 MAGIC
refering to `http://htmlpreview.github.io/?https://github.com/KrishnaswamyLab/MAGIC/blob/master/Rmagic/inst/examples/bonemarrow_tutorial.html`
## 1.0 Load data and packages
```{r load of packages and setup of the environment 1}
gc()
rm(list=ls())
library(Seurat)
library(SeuratWrappers)
library(sctransform)
library(harmony)
library(data.table)
library(scales)
library(plotly)
library(ggplot2)
library(ggpointdensity)
library(monocle3)
library(magrittr)
library(dplyr)
library(tidyverse)
library(cowplot)
library(clustree)
library(gridExtra)
library(phateR)
library(viridis)
library(readr)
library(Rmagic)

set.seed(12)
theme_set(theme_bw())

JE_combined_epi <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_20221128.rds")

```

Generate expression matrix

```{r}

JE_combined_epi_RNA<-GetAssayData(object = JE_combined_epi[["RNA"]], slot = "count")
JE_combined_epi_RNA_table<-as.data.frame(JE_combined_epi_RNA)

head(JE_combined_epi_RNA_table)
```
Transpose matrix
```{r}
JE_combined_epi_RNA_table_t <- as.matrix(t(JE_combined_epi_RNA_table))

```

```{r}
row.names(JE_combined_epi_RNA_table_t) <-NULL
JE_combined_epi_RNA_table_t[1:5,1:10]
```
Filter genes expression matrix
```{r}
# keep genes expressed in at least 10 cells
keep_cols <- colSums(JE_combined_epi_RNA_table_t > 0) > 10
JE_combined_epi_RNA_table_t <- JE_combined_epi_RNA_table_t[,keep_cols]
# look at the distribution of library sizes
ggplot() +
  geom_histogram(aes(x=rowSums(JE_combined_epi_RNA_table_t)), bins=50) +
  geom_vline(xintercept = 1000, color='red')
```
```{r}
# keep cells with at least 1000 UMIs
keep_rows <- rowSums(JE_combined_epi_RNA_table_t) > 1000
JE_combined_epi_RNA_table_t <- JE_combined_epi_RNA_table_t[keep_rows,]
```
Normalizing data

We should library size normalize and transform the data prior to MAGIC. Many people use a log transform, which requires adding a “pseudocount” to avoid log(0). We square root instead, which has a similar form but doesn’t suffer from instabilities at zero.
```{r}
JE_combined_epi_RNA_table_t <- library.size.normalize(JE_combined_epi_RNA_table_t)
JE_combined_epi_RNA_table_t <- sqrt(JE_combined_epi_RNA_table_t)
```






## 1.1 Further MAGIC for all genes
Using MAGIC for downstream analysis
First we will test from two mainly expressed genes
```{r}
JE_combined_epi_MAGIC <- magic(JE_combined_epi_RNA_table_t, genes=c("Krt14", "Krt5", "Axin2"))
```

We can look at the entire smoothed matrix with genes='all_genes', passing the original result to the argument init to avoid recomputing intermediate steps. Note that this matrix may be large and could take up a lot of memory.

```{r}
JE_combined_epi_MAGIC <- magic(JE_combined_epi, genes="all_genes",
                     t=4, init=JE_combined_epi_MAGIC)
```

```{r}
JE_combined_epi_MAGIC_RNA<-GetAssayData(object = JE_combined_epi_MAGIC[["MAGIC_SCT"]], slot = "data")
JE_combined_epi_MAGIC_RNA_table<-as.data.frame(JE_combined_epi_MAGIC_RNA)

head(JE_combined_epi_MAGIC_RNA_table)
```

This is a matrix for further correlation analysis of any genes of interest.

```{r}
JE_combined_epi_MAGIC_RNA_table_t <- t(JE_combined_epi_MAGIC_RNA_table)
JE_combined_epi_MAGIC_RNA_table_t[1:5,1:10]
```


```{r}
JE_combined_epi_MAGIC_RNA_table_t <- as.data.frame(JE_combined_epi_MAGIC_RNA_table_t)
ggplot(JE_combined_epi_MAGIC_RNA_table_t) +
  geom_point(aes(JE_combined_epi_MAGIC_RNA_table_t$Ggt1, JE_combined_epi_MAGIC_RNA_table_t$Odam, color=JE_combined_epi_MAGIC_RNA_table_t$Krt5)) +
  scale_color_viridis(option="B")
```

```{r}
ggplot(JE_combined_epi_MAGIC_RNA_table_t) +
  geom_point(aes(Moxd1, Axin2, color=Krt5)) +
  scale_color_viridis(option="B")
```

```{r}
ggplot(JE_combined_epi_MAGIC_RNA_table_t) +
  geom_point(aes(Bmi1, Axin2, color=Krt5)) +
  scale_color_viridis(option="B")
```
## 1.2 Standard Seurat for MAGIC object
*You may probe dims and resolution in details*
```{r, dimension reduction}
JE_combined_epi_MAGIC <- RunUMAP(object = JE_combined_epi_MAGIC, assay = "MAGIC_SCT", reduction = "harmony", dims = 1:50) #You may probe dims in details
Filter(f = function(x) inherits(JE_combined_epi_MAGIC[[x]], "Graph"), names(JE_combined_epi_MAGIC))
JE_combined_epi_MAGIC <- FindNeighbors(object = JE_combined_epi_MAGIC, assay = "MAGIC_SCT", reduction = "harmony", dims = 1:50) #You may probe dims in details
```
```{r}
Filter(f = function(x) inherits(JE_combined_epi_MAGIC[[x]], "Graph"), names(JE_combined_epi_MAGIC))
```


*For details of probing resolution, refer to `https://cloud.tencent.com/developer/article/1825681`. *
Here we will set an intermediate seurat object innheriting all the metadata from the formal object.
```{r, probing resolution parameter}

JE_combined_epi_MAGIC.resolution <- FindClusters(JE_combined_epi_MAGIC, dims=1:50,resolution = seq(from=0,by=.2,length=10))

```

```{r, fig.height=10, fig.width=10}
clustree(JE_combined_epi_MAGIC.resolution)
```
*Choosing resolution is quite arbitrary and no conclusion has been attained before biological validation*
The `suture.resolution` object has many choice for resultion, such as SCT_snn_res.0.8. You may want to plot and check. Here are some examples:
```{r, probeing resolution}
DimPlot(JE_combined_epi_MAGIC.resolution, reduction = "umap", label = T, group.by = "SCT_snn_res.0.2")
DimPlot(JE_combined_epi_MAGIC.resolution, reduction = "umap", label = T, group.by = "SCT_snn_res.0.4")
DimPlot(JE_combined_epi_MAGIC.resolution, reduction = "umap", label = T, group.by = "SCT_snn_res.0.6")
DimPlot(JE_combined_epi_MAGIC.resolution, reduction = "umap", label = T, group.by = "SCT_snn_res.0.8")

```

Here we choose res = 0.4


```{r, cluster with a suitable resolution}
JE_combined_epi_MAGIC <- FindClusters(object = JE_combined_epi_MAGIC, resolution = 0.4) #You may probe resolution in details
```

```{r, test plot,fig.height=5, fig.width=8}
plots <- DimPlot(JE_combined_epi_MAGIC, group.by = c("orig.ident", "seurat_clusters"), combine = FALSE, pt.size = .2)
plots <- lapply(X = plots, FUN = function(x) x + theme(legend.position = "top") + guides(color = guide_legend(nrow = 5, 
    byrow = TRUE, override.aes = list(size = 4))))
CombinePlots(plots)
```
## 1.3 Save object
```{r, output integrated data}
saveRDS(JE_combined_epi_MAGIC, file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_20221128.rds")
```



```{r, krt genes feature plot, fig.height=5, fig.width=5}
FeaturePlot(JE_combined_epi, features = c("Krt5","Lor",
                                          "Odam","Axin2"),
            cols = c("#ddf4f5", "#fa3c84"),
            ncol=2, pt.size = 0.1)
#ggsave("./OutputFigure/JE_combined_epi_krt_feature.pdf", device = "pdf", width = 15, height = 15, units = "cm")

```



```{r, krt genes feature plot, fig.height=5, fig.width=5}
FeaturePlot(JE_combined_epi_MAGIC, features = c("Krt5","Lor",
                                          "Odam","Top2a"),
            cols = c("#ddf4f5", "#fa3c84"), label = T,
            ncol=2, pt.size = 0.1)
#ggsave("./OutputFigure/JE_combined_epi_krt_feature.pdf", device = "pdf", width = 15, height = 15, units = "cm")

```
## 1.4 Rename cluster
According to our in vivo validation based on the marker genes in each cluster, we rename and merge each cluster accordingly
```{r}
JE_combined_epi_MAGIC <- readRDS("./SeuratObject/JE_combined_epi_harmony_MAGIC_20221128.rds")
```

```{r}
DimPlot(JE_combined_epi_MAGIC, reduction = "umap", split.by  = "orig.ident")
```


```{r}
DimPlot(JE_combined_epi_MAGIC, reduction = "umap", label = T, group.by = "SCT_snn_res.0.4")
```
cluster 9 and cluster 12 are noisy clusters, so we remove them arbitrarily.
```{r}
JE_combined_epi_MAGIC <- subset(JE_combined_epi_MAGIC, idents = c(0,1,2,3,4,5,6,7,8,10,11))
```

With GE stands for gingival epithelium, we renamed the following clusters.
```{r, rename cluster}
new.cluster.ids <- c("GE1", #Cluster 0 cells differentiating for wound re-epithelialization with marker genes
                     "GE2", #Cluster 1 cells differentiating for wound re-epithelialization with marker genes
                     "GE3", #Cluster 2 G1/S, G2/M cycling cells with marker genes
                     "GE4", #Cluster 3 Keratinocytes away from the wound edge with marker genes
                     "GE5", #Cluster 4 cells that immediate early response to wounding with marker genes
                     "GE6", #Cluster 5 cells with high metabolic level with marker genes
                     "GE7", #Cluster 6 Junctional epithelium with marker genes
                     "GE3", #Cluster 7 G1/S, G2/M cycling cells with marker genes
                     "GE8", #Cluster 8 cells with high metabolic level with marker genes
                     "9-removed", #Cluster 9 Keratinocytes away from the wound edge with marker genes
                     "GE9", #Cluster 10 Keratinocyte-fibroblasts with marker genes
                     "GE9", #Cluster 11 Keratinocyte-fibroblasts with marker genes
                     "10-removed" #Cluster 12 erythroid/other with marker genes
                     )
Idents(JE_combined_epi_MAGIC) <- c("SCT_snn_res.0.4")
JE_combined_epi_MAGIC@meta.data$seurat_clusters <-JE_combined_epi_MAGIC@meta.data$SCT_snn_res.0.4
JE_combined_epi_MAGIC@meta.data$celltype<- JE_combined_epi_MAGIC@meta.data$seurat_clusters
levels(JE_combined_epi_MAGIC@meta.data$celltype) <- new.cluster.ids
```

```{r}
DimPlot(JE_combined_epi_MAGIC, reduction = "umap", label = T, group.by = "celltype")
```

```{r, initial plots}
Idents(JE_combined_epi_MAGIC) <- c("celltype")
#group_by_cluster
plot1 = DimPlot(JE_combined_epi_MAGIC, reduction = "umap", label = T)
# group_by_sample
plot2 = DimPlot(JE_combined_epi_MAGIC, reduction = "umap", group.by = "celltype")
plot3 = DimPlot(JE_combined_epi_MAGIC, reduction = "umap", split.by  = "orig.ident")
# combine
plotc <- plot1 + plot2
ggsave("./OutputFigure/JE_combined_epi_MAGIC_umap.png", plot = plotc, width = 10, height = 5)
plot3
ggsave("./OutputFigure/JE_combined_epi_MAGIC_umap_split.png", plot = plot3, width = 12, height = 5) # You may set the parameters, especially the "width". Usually each plot will take about 5.5 unit.

DimPlot(JE_combined_epi_MAGIC, reduction = "umap", label = T)
DimPlot(JE_combined_epi_MAGIC, reduction = "umap", group.by = "celltype")

```

```{r, output integrated data}
saveRDS(JE_combined_epi_MAGIC, file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```
## 1.5 Marker genes for renamed cluster
```{r, generally find marker genes}
Idents(JE_combined_epi_MAGIC) <- JE_combined_epi_MAGIC$celltype
DefaultAssay(JE_combined_epi_MAGIC) <- "RNA"
JE_combined_epi_MAGIC <- NormalizeData(JE_combined_epi_MAGIC)
all.genes <- rownames(JE_combined_epi_MAGIC)
JE_combined_epi_MAGIC <- ScaleData(JE_combined_epi_MAGIC, features = all.genes)
JE_combined_epi_MAGIC_markers <- FindAllMarkers(object = JE_combined_epi_MAGIC, only.pos = TRUE, 
                               min.pct = 0.25, thresh.use = 0.25) #identify positive marker genes 
top10 <- JE_combined_epi_MAGIC_markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) # You may change n=10 to other number to get top_n genes as you wish.
```
```{r, output tables for marker genes}
write.csv(JE_combined_epi_MAGIC_markers, file = "./OutputTable/JE_combined_epi_MAGIC_markers_20221228.csv", row.names = FALSE)
write.csv(top10, file = "./OutputTable/JE_combined_epi_MAGIC_top10_markers_20221228.csv", row.names = FALSE)


```

Plot heatmap for marker genes
```{r, heatmapplot, fig.width=10, fig.height=18 }

DoHeatmap(JE_combined_epi_MAGIC, features = top10$gene)
ggsave("./OutputFigure/JE_combined_epi_MAGIC_markers_top10_20221228.pdf", device = "pdf",#adjust filename accordingly
       width = 30, #adjust according to the display of output
       height = 55, #adjust according to the display of output
       units = "cm")

```

# 2 3D output for UMAP
```{r}
JE_combined_epi_MAGIC <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```
Here we redo the dimension reduction from three layers.
```{r}
JE_combined_epi_MAGIC <- RunUMAP(object = JE_combined_epi_MAGIC, assay = "MAGIC_SCT", reduction = "harmony", dims = 1:50,
                                 n.components = 3L) #You may probe dims in details
```
```{r}
tmp.umap.3<-Embeddings(object = JE_combined_epi_MAGIC[["umap"]])
```


```{r}
cb_palette <- c("#F8766D", #JE1
                "#D39200", #JE2
                "#93AA00", #JE3
                "#00BA38", #JE4
                "#00C19F", #JE5
                "#00B9E3", #JE6
                "#619CFF", #JE7
                "#DB72FB", #JE8
                "#FF61C3" #JE9
                )
               
```

## 2.1 3D UMAP 
```{r}
library(plotly)
tmp.umap.3 <- as.data.frame(tmp.umap.3)
fig <- plot_ly(tmp.umap.3, x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, color =JE_combined_epi_MAGIC$celltype, 
               colors = cb_palette,
               opacity = .5,
               type = "scatter3d", 
               size=1)
fig
```
From the 3D plot, we can find GE7 (Odam+ cluster) is tightly connected with GE5 and GE9

## 2.2 3D genes of interest
```{r, 3D plot for Odam}
# create a dataframe
goi <- "Odam"
plotting.data <- FetchData(object = JE_combined_epi_MAGIC, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .99)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called Odam
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~Odam, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c("#ddf4f5", "#fa3c84"), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 2), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)
```

```{r, 3D plot for Cxcl5}
# create a dataframe
goi <- "Cxcl5"
plotting.data <- FetchData(object = JE_combined_epi_MAGIC, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .99)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called Cxcl5
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~Odam, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c("#ddf4f5", "#fa3c84"), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 2), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)
```

```{r, 3D plot for Axin2}
# create a dataframe
goi <- "Axin2"
plotting.data <- FetchData(object = JE_combined_epi_MAGIC, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .99)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called Axin2
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~Axin2, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c("#ddf4f5", "#fa3c84"), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 2), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)
```
```{r, 3D plot for Top2a}
# create a dataframe
goi <- "Top2a"
plotting.data <- FetchData(object = JE_combined_epi_MAGIC, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .99)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called ACTB
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~ACTB, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c("#ddf4f5", "#fa3c84"), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 2), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)
```
```{r, 3D plot for Lor}
# create a dataframe
goi <- "Lor"
plotting.data <- FetchData(object = JE_combined_epi_MAGIC, vars = c("UMAP_1", "UMAP_2", "UMAP_3", "Expression"=goi), slot = 'data')

# Say you want change the scale, so that every cell having an expression >1 will be one color
# Basically, you are re-adjusting the scale here, so that any cell having a certain expression will light up on your 3D plot

# First make another column in your dataframe, where all values above 1 are re-assigned a value of 1
# This information is stored in the 'Expression' column of your dataframe
# Cutoff <- 2
Cutoff <- quantile(plotting.data[,goi], probs = .99)
plotting.data$"ExprCutoff" <- ifelse(test = plotting.data[,goi] <Cutoff, yes = plotting.data[,goi], no = Cutoff)

# Add the label column, so that now the column has 'cellname-its expression value'
plotting.data$label <- paste(rownames(plotting.data)," - ", plotting.data[,goi], sep="")

# Plot your data, in this example my Seurat object had 21 clusters (0-20), and cells express a gene called Lor
plot_ly(data = plotting.data,
        # name = goi,
        x = ~UMAP_1, y = ~UMAP_2, z = ~UMAP_3, 
        color = ~ExprCutoff, # you can just run this against the column for the gene as well using ~Lor, the algorith will automatically scale in that case based on maximal and minimal values
        opacity = .5,
        colors = c("#ddf4f5", "#fa3c84"), 
        type = "scatter3d", 
        mode = "markers",
        marker = list(size = 2), 
        text=~label,
        hoverinfo="text"
) %>%layout(title=goi)
```




## 2.3 Save the output
```{r}
saveRDS(JE_combined_epi_MAGIC, "./SeuratObject/JE_combined_epi_MAGIC_3D.rds")

```





#3 Generate scVelo attributes
Load data again.
```{r}
JE_combined_epi_MAGIC <- readRDS(file = "./SeuratObject/JE_combined_epi_harmony_MAGIC_renamed.rds")
```

```{r, prepare parameter matrix of combined JE for scvelo}
# Export filtered cell ID
JE_combined_epi_MAGIC_cells <- as.data.frame(Cells(JE_combined_epi_MAGIC))
head(JE_combined_epi_MAGIC_cells)
write.csv(JE_combined_epi_MAGIC_cells, file = "./scVeloInput/JE_combined_epi_MAGIC_cellID_obs.csv",row.names = FALSE)

# Export umap coordinates
umap_JE_combined_epi_MAGIC <- Embeddings(JE_combined_epi_MAGIC, reduction = "umap")
head(JE_combined_epi_MAGIC)
umap_JE_combined_epi_MAGIC[1:3,1:2]
write.csv(umap_JE_combined_epi_MAGIC, file = "./scVeloInput/JE_combined_epi_MAGIC_cell_embeddings.csv",row.names = TRUE)

# Export cluster information
## Here we used celltype.
head(JE_combined_epi_MAGIC@meta.data$celltype) #
cluster_head <-JE_combined_epi_MAGIC@meta.data$celltype # Optional
cluster_head <- as.data.frame(cluster_head)
cluster_info <- cbind(JE_combined_epi_MAGIC_cells$`Cells(JE_combined_epi_MAGIC)`, cluster_head)
colnames(cluster_info) <- c("Unnamed", "celltype") #if you are working with celltype, change `seurat_cluster` into `celltype` accordingly.
write.csv(cluster_info, file = "./scVeloInput/JE_combined_epi_MAGIC_clusters.csv",row.names = FALSE)
```


# 4 Subset two paths
From PAGA analysis, we can clear find there are two paths: GE1, GE5, GE9, GE3 and GE7; GE1 GE8 GE2, GE6 and GE4
We are going to subset these two paths for monocle analysis
```{r}
JE_combined_epi_MAGIC_GE7 <- subset(JE_combined_epi_MAGIC, idents = c("GE1", "GE5", "GE9", "GE3", "GE7"))

JE_combined_epi_MAGIC_GE4 <- subset(JE_combined_epi_MAGIC, idents = c("GE1", "GE8", "GE2", "GE6", "GE4"))
```


```{r}
DimPlot(JE_combined_epi_MAGIC_GE7, reduction = "umap", label = T)
DimPlot(JE_combined_epi_MAGIC_GE7, reduction = "umap", group.by = "celltype")

```


```{r}
DimPlot(JE_combined_epi_MAGIC_GE4, reduction = "umap", label = T)
DimPlot(JE_combined_epi_MAGIC_GE4, reduction = "umap", group.by = "celltype")

```

```{r}
saveRDS(JE_combined_epi_MAGIC_GE7, file = "./SeuratObject/JE_combined_epi_MAGIC_GE7_20221128.rds")
saveRDS(JE_combined_epi_MAGIC_GE4, file = "./SeuratObject/JE_combined_epi_MAGIC_GE4_20221128.rds")
```

